---
title: "VBM Evaluation"
author: "Lucy Kim, Simon Jones"
date: "3/19/2018"
output: html_document
---

# Evaluation of the VBM

## Motivation
The NYU Langone Health has been consistently ranked high in terms of quality but also high in terms of costs per case among academic medical centers nationally even after the adjustment for the case mix and wage indices. Thus, the NYULMH implemented value-based management program to achieve reducing costs and improving value of care which can be defined as quality achieved per dollar spent. We aim to evaluate the impact of this program in terms of variable direct cost per case and quality, and estimate cost savings to date.

This analysis code marks May 2014 as the first month of the post-intervention period. 

## Data set up
```{r}
rm(list = ls())

library(data.table)
library(stats)
require(forecast)
library(forcats)
library(nlme)
library(readxl)
library(stringr)
library(sqldf)
library(broom)
library(stargazer)
library(lubridate)
library(lme4)
library(Hmisc)
library(plyr)
library(doBy)
library(clusterSEs)
library(censReg)
library(qcc)
library(ggplot2)
library(dplyr)
library(MASS)
library(reshape2)
library(scales)

load("~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")
# load("~/Box Sync/VBM/Data/sample-full2017-raw.RData")

# pat <- read_excel("~/Box Sync/VBM/Data/Patient Level Data with expense grouping deidentified 1052017.xlsx", guess_max=288364)
# pat2 <- read_excel("~/Box Sync/VBM/Data/Sep 2016 - Dec 2017 VBM Paper deintentified.xlsx", guess_max=94160)

NROW(pat)
# str(pat)
pat$AdmitDateD   <- as.Date( pat$`Admit Date` )
pat$DischargeDateD   <- as.Date( pat$`Discharge Date` )

min(pat$DischargeDateD ) # "2010-09-01"
max(pat$DischargeDateD ) #  "2017-05-31"

NROW(pat2)
# str(pat2)
pat2$AdmitDateD   <- as.Date( pat2$`Admit Date` ) 
pat2$DischargeDateD   <- as.Date( pat2$`Discharge Date` )

min(pat2$DischargeDateD ) #  "2016-09-01"
max(pat2$DischargeDateD ) #  "2017-12-31"

# recode race categories in the new discharge data for FY 2017 using the xwalk 
racexwalk <- read.csv("~/Box Sync/VBM/Data/race_xwalk.csv")
test <- merge(pat2, racexwalk, by.x="Race", by.y="alphabet", all.x = T)
test$Race <- test$race_des
pat2 <- test
rm(test)

#subset the first data to Sep 2011 - Aug 2016 (exclude FY 11 & 2017)
pat3 <- subset(pat, pat$DischargeDateD >= "2011-09-01" & pat$DischargeDateD < "2016-09-01")
min(pat3$DischargeDateD ) # "2011-09-01"
max(pat3$DischargeDateD ) #  "2016-08-31"

# before appending data, remove problematic variables
v <- names(pat3) %in% c("NYU Reporting Department – Most Recent", "NYU Reporting Division – Most Recent")
pat3 <- pat3[!v]
v <- names(pat2) %in% c("NYU Reporting Department – Most Recent", "NYU Reporting Division – Most Recent")
pat2 <- pat2[!v]
pat3$`Admit Service` <- as.numeric(pat3$`Admit Service`)
pat3$`Admit Source` <- as.numeric(pat3$`Admit Source`)
pat3$`Admission Type` <- as.numeric(pat3$`Admission Type`)
pat3$`EPIC Discharge Dept` <- as.numeric(pat3$`EPIC Discharge Dept`)
pat3$`Discharge Service` <- as.numeric(pat3$`Discharge Service`)
pat3$`Discharge Disposition` <- as.numeric(pat3$`Discharge Disposition` )
pat3$`MS-DRG` <- as.numeric(pat3$`MS-DRG`)
pat3$`NY-DRG` <- as.numeric(pat3$`NY-DRG`)

# sum(is.na(pat3$`Enc - Primary ICD10 Surgical Procedure`))
# pat3$`Enc - Primary ICD10 Surgical Procedure` <- as.character(pat3$`Enc - Primary ICD10 Surgical Procedure`)
# # pat3 %<>% mutate_all(funs(ifelse(is.na(.)&is.numeric(.),NA,.)))
# sum(is.na(pat3$`NYU Vizient SSI Infection Flag`))
# pat3$`NYU Vizient SSI Infection Flag` <- as.character(pat3$`NYU Vizient SSI Infection Flag`)

# append the two datasets
patUse <- bind_rows(pat3, pat2, .id="id" )
table(patUse$id)
min(patUse$DischargeDateD ) # "2011-09-01"
max(patUse$DischargeDateD ) #  "2017-12-31"

patUse$DischargeDateMonth <- format(patUse$DischargeDateD  , "%Y%m")  
NROW(patUse)
# test <- subset(patUse, !is.na(`Enc - Primary ICD10 Surgical Procedure`))
# View(test$`Enc - Primary ICD10 Surgical Procedure`)
```

Quick data summary
```{r}
# discharge department
table(patUse$`EPIC Discharge Dept_Desc`)
table(patUse$DischargeDateMonth , is.na(patUse$`EPIC Discharge Dept_Desc`))

# check when Lutheran appears
library(stringr)
table(patUse$DischargeDateMonth , str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) == "L") # starts in 201608
table(patUse$DischargeDateMonth , str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) == "X") # appears only in 2011/09 - 2015/08

# get mean of every numeric columns by month
test <- ddply(patUse, .(DischargeDateMonth), numcolwise(mean))
# expected LOS = 0 before 2012/09
```


Exclude non-inpatient people
```{r}
sum(is.na(patUse$PatientClass_Desc))
table(patUse$PatientClass_Desc)
patUse <- subset(patUse, PatientClass_Desc=="Inpatient")
NROW(patUse)
```


exclude DRG weight = 0 or MS-DRG = 998 or 999
```{r}
# drop MS-DRG = 999
summary(patUse$`MS DRG Weight (CMI)`)
patUse <- subset(patUse, `MS-DRG`!=999 & `MS-DRG`!=998 & `MS DRG Weight (CMI)`!=0)
summary(patUse$`MS DRG Weight (CMI)`)
NROW(patUse)
```


Exclude Rehab hospital & rehab discharges by using reporting department = Rehab or Psych (these are not typically paid using DRG)
```{r}
# patUse <- patUse[(str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) != "X" & str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) != "L" )| is.na(patUse$`EPIC Discharge Dept_Desc`) , ]
# patUse <- subset(patUse, `MS-DRG Type`!="REHAB")

# exclude MS-DRG type is missing and reporting department is missing
patUse <- patUse[complete.cases(patUse[, c("NYU Reporting Department – Most Recent_Desc", "MS-DRG Type")]),]

# exclude if reporting department is unknown
patUse <- subset(patUse, `NYU Reporting Department – Most Recent_Desc`!="UNK")

# rehab dept can use non-rehab diagnosis
table(patUse$`NYU Reporting Department – Most Recent_Desc`, patUse$`MS-DRG`==945 | patUse$`MS-DRG`==946)

# distribution of medical, surgical, rehab patients by reporting dept
table(patUse$`NYU Reporting Department – Most Recent_Desc`, patUse$`MS-DRG Type`)

# define rehab as Report dept=Rehab & MS DRG=945/946
# patUse$pattype <- factor(ifelse( patUse$`NYU Reporting Department – Most Recent_Desc`=="Rehabilitation Medicine", 1, ifelse((patUse$`MS-DRG`==945 | patUse$`MS-DRG`==946) & patUse$`NYU Reporting Department – Most Recent_Desc`=="Rehabilitation Medicine", 2, ifelse(patUse$`MS-DRG Type`=="MED", 3, ifelse(patUse$`MS-DRG Type`=="SURG", 4, 5))), labels=c("Report dept=Rehab", "Report dept=Rehab & MS DRG=945/946", "Medical", "Surgical", "Unknown")))

patUse$pattype <- factor(ifelse( (patUse$`MS-DRG`==945 | patUse$`MS-DRG`==946) & patUse$`NYU Reporting Department – Most Recent_Desc`=="Rehabilitation Medicine", 1, ifelse( patUse$`NYU Reporting Department – Most Recent_Desc`=="Psychiatry", 2, ifelse(patUse$`MS-DRG Type`=="MED", 3, ifelse(patUse$`MS-DRG Type`=="SURG", 4, 5)))), labels=c("Report dept=Rehab & MS DRG=945/946", "Report dept=Psych", "Medical", "Surgical", "Unknown"))
table(patUse$pattype)

# plot # medical, surgical, rehab discharges in each month?
test <- ddply(patUse, .(DischargeDateMonth, pattype), summarise, N=length(pattype))
test$N[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA

ggplot(test, aes(x=DischargeDateMonth, y=N, colour=pattype, group=pattype)) +
  geom_line(base_size=10) +
  labs(x="Month", y="Number of discharges",
       title="Number of medical, surgical, rehab patient discharges by month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  geom_vline(aes(xintercept=50), colour="blue") +
  scale_y_continuous(limits = c(0,2400),
                     breaks=seq(0,2400,200))
ggsave("~/Dropbox/Research/VBM/results/num_discharges1.png", width=20, height=10, units="cm")

# define rehab as Report dept=Rehab only
patUse$pattype <- factor(ifelse( patUse$`NYU Reporting Department – Most Recent_Desc`=="Rehabilitation Medicine", 1, ifelse(patUse$`MS-DRG Type`=="MED", 2, ifelse(patUse$`MS-DRG Type`=="SURG", 3, 4))), labels=c("Report dept=Rehab", "Medical", "Surgical", "Unknown"))

# plot # medical, surgical, rehab discharges in each month?
test <- ddply(patUse, .(DischargeDateMonth, pattype), summarise, N=length(pattype))
test$N[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA

ggplot(test, aes(x=DischargeDateMonth, y=N, colour=pattype, group=pattype)) +
  geom_line() +
  labs(x="Month", y="Number of discharges",
       title="Number of medical, surgical, rehab patient discharges by month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=33), colour="red") +
  geom_vline(aes(xintercept=50), colour="blue") +
  scale_y_continuous(limits = c(0,2400),
                     breaks=seq(0,2400,200))
ggsave("~/Dropbox/Research/VBM/results/num_discharges2.png", width=20, height=10, units="cm")

# since unknown patients stop after 2014/07, drop these patients - they have MS-DRG = Rehab
test2 <- subset(patUse, `NYU Reporting Department – Most Recent_Desc`!="Rehabilitation Medicine" & `MS-DRG Type`!="MED" & `MS-DRG Type`!="SURG")
table(test2$DischargeDateMonth, test2$`MS-DRG Type`) # all MS-DRG Type = rehab

# exclude MS-DRG Type = rehab & reporting dept = "Rehab medicine"
patUse <- subset(patUse, `NYU Reporting Department – Most Recent_Desc`!="Rehabilitation Medicine" & `MS-DRG Type`!="REHAB")
NROW(patUse)

# exclude psych patients
patUse <- subset(patUse, patUse$`NYU Reporting Department – Most Recent_Desc`!="Psychiatry")
NROW(patUse)

# re-plot # discharges by patient type across months
test <- ddply(patUse, .(DischargeDateMonth, pattype), summarise, N=length(pattype))
test$N[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA

ggplot(test, aes(x=DischargeDateMonth, y=N, colour=pattype, group=pattype)) +
  geom_line() +
  theme_bw(base_size=12) +
  labs(x="Month", y="Number of discharges",
       title="Number of medical, surgical patient discharges by month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  scale_y_continuous(limits = c(0,2400),
                     breaks=seq(0,2400,200))
ggsave("~/Dropbox/Research/VBM/results/num_discharges3.png", width=20, height=10, units="cm")


# these 4 places were for rehab patients
# test <- subset(patUse, (`EPIC Discharge Dept_Desc` %in% c("HJD 8 SOUTH", "HJD 9 NORTH", "HJD 9 SOUTH", "TH HCC 9")))
# table(test$DischargeDateMonth)
# table(test$`Discharge Service_Desc`)

# drop those 4 places
# patUse <- subset(patUse, !(`EPIC Discharge Dept_Desc` %in% c("HJD 8 SOUTH", "HJD 9 NORTH", "HJD 9 SOUTH", "TH HCC 9")))
table(patUse$`Discharge Service_Desc`)

# drop any patients whose discharge service dept is Rehab medicine or Rehabilitation
# patUse <- subset(patUse, !(`Discharge Service_Desc` %in% c("REHAB MEDICINE", "Rehabilitation")))

table(patUse$DischargeDateMonth)
```

Exclude newborns (MS-DRG = 795); these can impact calculations when volume fluctuates for newborns as the mother and newborns are both counted
```{r}
table(patUse$`MS-DRG`)
patUse <- subset(patUse, `MS-DRG`!=795)
NROW(patUse)
```

Exclude Lutheran hospitals because they joined the hospital system later  
```{r}
# exclude Lutheran
patUse <- subset(patUse, str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) != "L" | is.na(patUse$`EPIC Discharge Dept_Desc`))
NROW(patUse)

table(patUse$DischargeDateMonth)

# focus on inpatient cases in Tisch 
table(patUse$`Custom Patient Type`)
table(patUse$DischargeDateMonth, patUse$`Custom Patient Type`)
# HI patients : ~500 patients per month
patUse <- subset(patUse, !is.na(`Custom Patient Type`) & `Custom Patient Type` %in% c("TI"))
#  "RI"
NROW(patUse)
```


Exclude LOS = 1 day and reporting department = Medicine - Don't apply this rule any more
```{r}
sum(is.na(patUse$`NYU Reporting Department – Most Recent_Desc`))
table(patUse$`NYU Reporting Department – Most Recent_Desc`) # "Medicine"

patUse$los_eq1 <- ifelse(patUse$`Length of Stay`==1 & patUse$`NYU Reporting Department – Most Recent_Desc`=="Medicine", 1,0)
table(patUse$los_eq1)

# test <- subset(patUse, patUse$`NYU Reporting Department – Most Recent_Desc`=="Medicine")
# test <- ddply(test, .(DischargeDateMonth), summarise, share = mean(los_eq1), num = sum(los_eq1))
# 
# test$num[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA
# test$share[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA
# 
# # plot the share of discharges who have LOS=1 & reporting dept = medicine
# ggplot(test, aes(x=DischargeDateMonth, y=share, group=1)) +
#   geom_line() +
#   ylim(0,1) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#   geom_vline(aes(xintercept=33), colour="red") +
#   labs(title="Share of discharges having LOS=1 in the reporting department Medicine",
#        x="Month", y="Share of discharges in the Medicine department")
# ggsave("~/Dropbox/Research/VBM/results/share_los1_medicine.png", width=20, height=10, units="cm")
# 
# ggplot(test, aes(x=DischargeDateMonth, y=num, group=1)) +
#   geom_line() +
#   # ylim(0,1) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#   geom_vline(aes(xintercept=33), colour="red") +
#   labs(title="Number of discharges having LOS=1 in the reporting department Medicine",
#        x="Month", y="Number of discharges in the Medicine department")
# ggsave("~/Dropbox/Research/VBM/results/num_los1_medicine.png", width=20, height=10, units="cm")
# 
# patUse <- subset(patUse, los_eq1==0)
```


## Adjust costs for inflation
Before anything, Robustness Check: Compare with the stats reported in "VBM Task Force Deck_Aug 2017 (DRAFT).pdf"

add monthly health care inflation data from BLS
```{r}
infl <- read_excel("~/Box Sync/VBM/Data/medical_inflation_03102018.xlsx", range = cell_rows(12:132), col_names=TRUE) # seasonally adjusted

# create month 
infl$day <- rep(1,nrow(infl))
infl$Period <-gsub('M','', infl$Period)
infl$Period <- as.numeric(infl$Period)
infl$date <- paste(infl$Year, infl$Period, infl$day, sep="/")
infl$date <- as.Date(infl$date)
infl$month <- format(infl$date, "%Y%m")
  
colnames(infl)[4] <- "cpi"

infl <- subset(infl, date >= "2011-09-01" & date <= "2017-12-01") # keep 2010/09 - 2017/05

infl <- infl[c("month","cpi")]

infl.ts <- ts(infl$cpi, start=c(2011,9), end=c(2017,12), frequency = 12)
plot(infl.ts, ylab="CPI", ylim=c(0,500))
dev.off()

# merge with discharge-level data
patUse <- merge(patUse, infl, by.x="DischargeDateMonth", by.y="month")
table(patUse$DischargeDateMonth)
```

adjust the costs for inflation using the base period = 2010/09 (first month in data)
```{r}
colnames(patUse)[which(colnames(patUse)=="Enc - Total Variable Direct Cost")] <- "tvdc"
colnames(patUse)[which(colnames(patUse)=="ICU/CCU")] <- "ICU"
colnames(patUse)[which(colnames(patUse)=="Operating Room")] <- "OR"
colnames(patUse)[which(colnames(patUse)=="Medical/Surgical Supplies")] <- "supplies"
colnames(patUse)[which(colnames(patUse)=="Inhalation Therapy")] <- "inhalation"
colnames(patUse)[which(colnames(patUse)=="Physical/Occupational Therapy")] <- "POT"
colnames(patUse)[which(colnames(patUse)=="Post Op/Stepdown")] <- "postop"
colnames(patUse)[which(colnames(patUse)=="Routine Care")] <- "routine"

costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "Implants","Blood", "ICU", "OR", "supplies", "inhalation", "POT", "postop", "routine")

# rebase using 2011/09 CPI
for (col in costs) {
  new.col <- paste(col, '_cpi', sep='')
  patUse[, new.col] <- patUse[, col] * 401.605/patUse$cpi
}

# check that inflation adjustment was done well
monthly_cost <- ddply(patUse, .(DischargeDateMonth), summarise, tvdcm=mean(tvdc, na.rm=TRUE), tvdc_cpim=mean(tvdc_cpi, na.rm=TRUE))

# create monthly time-series data
monthly_cost.ts <- ts(monthly_cost, start = c(2011,9), end=c(2017,5), frequency = 12)
# monthly_cost.ts <- ts(monthly_cost[25:nrow(monthly_cost),], start = c(2013,9), end=c(2017,5), frequency = 12)

pdf("~/Dropbox/Research/VBM/results/inflation_compare.pdf")
par(mfrow=c(2,1))
ts.plot(infl.ts, ylab="CPI", ylim=c(0,500))
ts.plot(monthly_cost.ts[,2:3],
        gpars=list(xlab="Month", ylab="Avg Total Variable Direct Cost", lty=c(1:2)),
        ylim=c(0,20000))
legend("topright", legend = c("Unadjusted", "CPI-adjusted"), lty = 1:2, cex=.65)
dev.off()
```

### education: compare mean monthly CMI adjusted VDC per inpatinet case (p.9) = mean (TVDC / CMI) for each month (skip for analysis)
```{r}
test <- ddply(patUse, .(DischargeDateMonth), summarise, mcosts = mean(tvdc, na.rm=TRUE), mcmi = mean(`MS DRG Weight (CMI)`, na.rm=TRUE), mcosts_cpi = mean(tvdc_cpi, na.rm=TRUE))
test$tvdc_adj1 <- test$mcosts / test$mcmi
test$tvdc_cpi_adj1 <- test$mcosts_cpi / test$mcmi

ggplot(test[test$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=tvdc_adj1)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(tvdc_adj1, digits = 2)), angle = 90, hjust=0)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0,8000) +
  geom_vline(aes(xintercept=9), colour="red") +
  labs(y="Mean total variable direct costs / Mean CMI")
ggsave("~/Dropbox/Research/VBM/results/CMIchng_theirway.pdf", width=20, height=10, units="cm")
# mean Costs / mean CMI

ggplot(test[test$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=tvdc_cpi_adj1)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(tvdc_cpi_adj1, digits = 2)), angle = 90, hjust=0)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0,8000) +
  geom_vline(aes(xintercept=9), colour="red") +
  labs(y="Mean CPI-adj total variable direct costs / Mean CMI")
ggsave("~/Dropbox/Research/VBM/results/CMIchng_theirway_cpi.pdf", width=20, height=10, units="cm")
# mean Costs / mean CMI

patUse$adj <- patUse$tvdc / patUse$`MS DRG Weight (CMI)`
patUse$adj_cpi <- patUse$tvdc_cpi / patUse$`MS DRG Weight (CMI)`
test2 <- ddply(patUse[patUse$`MS DRG Weight (CMI)` > 0,], .(DischargeDateMonth), summarise, tvdc_adj1 = mean(adj, na.rm=TRUE), tvdc_cpi_adj1 = mean(adj_cpi, na.rm=TRUE))

ggplot(test2[test2$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=tvdc_adj1)) +
  # geom_bar(stat="summary", fun.y="mean") +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(tvdc_adj1, digits = 2)), angle = 90, hjust=0)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0,8000) +
    geom_vline(aes(xintercept=9), colour="red") +
  labs(y="Mean [total variable direct costs / CMI]")
ggsave("~/Dropbox/Research/VBM/results/CMIchng_aggmean.pdf", width=20, height=10, units="cm")

ggplot(test2[test2$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=tvdc_cpi_adj1)) +
  # geom_bar(stat="summary", fun.y="mean") +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(tvdc_cpi_adj1, digits = 2)), angle = 90, hjust=0)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0,8000) +
  labs(y="Mean [CPI-adj total variable direct costs / CMI]") +
  geom_vline(aes(xintercept=9), colour="red")
ggsave("~/Dropbox/Research/VBM/results/CMIchng_aggmean_cpi.pdf", width=20, height=10, units="cm")
# mean Costs / mean CMI

# linear <- lm(tvdc_cpi ~ 0 + DischargeDateMonth + CMI,
#    data=patUse,
#    na.action="na.omit")
#
# summary(linear)
```

### education: compare mean blood, labs, ICU/CCU, routine care, post op/stepdown costs (skip for analysis)
```{r}
# compare only fy14 sep-feb vs fy17 dec-may
test <- subset(patUse, (DischargeDateMonth >= "201309" & DischargeDateMonth <= "201402") | (DischargeDateMonth >= "201612" & DischargeDateMonth <= "201702") )

test$current <- factor(ifelse(test$DischargeDateMonth >= "201309" & test$DischargeDateMonth <= "201402",0,1), labels = c("Baseline","Current"))

test2 <- ddply(test, .(current), summarise, Blood = mean(Blood, na.rm=TRUE))

ggplot(test2, aes(x=current, y=Blood)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(Blood, digits = 2)), vjust=0)) +
  labs(x="", y=paste0("Blood"," ($)"))

# plot monthly
patUse$current <- factor(ifelse(patUse$DischargeDateMonth >= "201309" & patUse$DischargeDateMonth <= "201402",1,ifelse(patUse$DischargeDateMonth >= "201612" & patUse$DischargeDateMonth <= "201702", 2, 3)), labels = c("Baseline","Current", "Other"))

costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "Implants","Blood", "ICU", "OR", "supplies", "inhalation","POT", "postop", "routine")
costs_cpi <- paste(costs, "_cpi", sep="")

test3 <- ddply(patUse, .(DischargeDateMonth, current), summarise, Blood = mean(Blood, na.rm=TRUE), Blood_cpi = mean(Blood_cpi, na.rm=TRUE))
test3$Blood[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308"] <- NA
test3$Blood_cpi[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308"] <- NA

ggplot(test3, aes(x=DischargeDateMonth, y=Blood_cpi, fill=current)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=sprintf("%0.2f", round(Blood_cpi, digits = 2)), vjust=0)) +
  labs(x="", y=paste0("CPI-adjusted", "Blood"," ($)")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
    geom_vline(aes(xintercept=45), colour="red")
ggsave("~/Dropbox/Research/VBM/results/blood_monthly.pdf", width=20, height=10, units="cm")

# still 2011/05-2011/09 have such low mean blood costs - espcially 2011/06 - 2011/08
test3[test3$DischargeDateMonth >= "201105" & test3$DischargeDateMonth <= "201109", c('Blood_cpi')]

# get monthly desc stats
test <- subset(patUse, DischargeDateMonth >= "201105" & DischargeDateMonth <= "201109")

table(patUse$DischargeDateMonth, patUse$Blood_cpi==0)
# zero blood cost obs counts are similar across months

ggplot(test, aes(Blood_cpi, colour=DischargeDateMonth)) +
  geom_density() +
  coord_cartesian(xlim=c(0, 5000))

ddply(test[test$Blood_cpi > 0,], .(DischargeDateMonth), summarise,
      min=min(Blood_cpi, na.rm=TRUE),
      M=mean(Blood_cpi, na.rm=TRUE),
      max=max(Blood_cpi, na.rm=TRUE),
      p25=quantile(Blood_cpi, p=(0.25), na.rm=TRUE),
      p50=quantile(Blood_cpi, p=(0.5), na.rm=TRUE),
      p75=quantile(Blood_cpi, p=(0.75), na.rm=TRUE)
)
# just the blood cost distribution is much more skewed towards zero during 2011/06-2011/08
```


## Create patient controls

Age group dummies
```{r}
table(patUse$Age)
sum(is.na(patUse$Age))==0

# quciker way of dping the below
patUse$AgeGroup <- cut(patUse$Age,breaks = c(seq(0,90, 5), 130), include.lowest = T)
```

payor group controls
```{r}
table(patUse$`NYU Finance - Payor Group`)
patUse$ins <- patUse$`NYU Finance - Payor Group`
sum(is.na(patUse$ins))

# Medicare, MA, T-Medicaid, Managed Medicaid, Self-pay, private
patUse$ins <- factor(ifelse(patUse$ins=="Medicare", 1, ifelse(patUse$ins=="Medicare HMO", 2, ifelse(patUse$ins=="Medicaid", 3, ifelse(patUse$ins=="Medicaid HMO", 4, ifelse(patUse$ins=="Self Pay", 5, 6))))), labels = c("Medicare FFS", "Medicare MC", "Medicaid FFS", "Medicaid MC", "Self-pay", "Commercial/Other ins"))
```

Gender
```{r}
table(patUse$Gender)
sum(is.na(patUse$Gender))==0
patUse$female <- factor(ifelse(patUse$Gender=="Female" | patUse$Gender=="F",1,0))
```

Race
```{r}
table(patUse$Race, patUse$id)

sum(is.na(patUse$Race))==0 # false; 42 obs have missing Race

patUse$raceGroup <- factor(ifelse(patUse$Race=="White",1,ifelse(patUse$Race=="African American (Black)",2, ifelse(patUse$Race!="Patient Refused" & patUse$Race!="" & patUse$Race!="Unknown",3,4))), labels = c("White", "Black", "Other race", "Unknown"))
patUse$raceGroup[is.na(patUse$Race)] <- "Unknown"
table(patUse$raceGroup)
```

seasonal FE
```{r}
patUse$month <- month(patUse$DischargeDateD)
patUse$season <- factor(ifelse(patUse$month>=3 & patUse$month<=5, 1, ifelse(patUse$month>=6 & patUse$month<=8, 2, ifelse(patUse$month>=9 & patUse$month<=11, 3, 4))), labels = c("Spring", "Summer", "Fall", "Winter"))
```

dummy for surgical/medical/rehab patient
```{r}
table(patUse$`MS-DRG Type`)
colnames(patUse)[which(colnames(patUse)=="MS-DRG Type")] <- "surgical"
patUse$surgical <- factor(ifelse(patUse$surgical=="SURG", 1, 0))
```

CMI
```{r}
colnames(patUse)[which(colnames(patUse)=="MS DRG Weight (CMI)")] <- "CMI"
```

Site indicator
```{r}
table(str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1))

patUse$site <- str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1)
```

Strategic area
```{r}
colnames(patUse)[which(colnames(patUse)=="Strategic Area")] <- "area"
colnames(patUse)[which(colnames(patUse)=="Strategic Area II")] <- "area2"

patUse$area[patUse$area=="Cardio & Vascular"] <- "Cardio Vascular"
patUse$area[patUse$area2=="Cardio & Vascular"] <- "Cardio Vascular"

patUse$area[patUse$area=="Neonatology"] <- "Children Services"
patUse$area[patUse$area=="Others"] <- "Other (Med)"
```

create time trend, post dummmy, time post intervention trend variables
```{r}
allDates <- sort(unique(patUse$DischargeDateMonth))

lastpremonthidx <- 32 # 36 for Aug 2014
nmonths <- length(allDates)

# will drop April 2014 - Aug 2014 later as this is a transition period (first VBM mtg on 4/19/2014)

dateLookUp <- data.frame(dayMonth = allDates,
                         t = 1:nmonths,
                         int = c(rep(0, lastpremonthidx), rep(1, nmonths-lastpremonthidx)),
                         tpostInt = c(rep(0, lastpremonthidx), 1:(nmonths-lastpremonthidx)))

patUse <- merge(patUse, dateLookUp, by.x="DischargeDateMonth", by.y="dayMonth")

# check the indexing was well done
test <- ddply(patUse, .(DischargeDateMonth), summarise, t = mean(t, na.rm=TRUE), int = mean(int, na.rm=TRUE), tpostInt = mean(tpostInt, na.rm=TRUE))
View(test)
```

create intensity score for supply & pharmacy
```{r}
intensity <- read_excel("~/Box Sync/VBM/Data/2017 Supply and Pharmacy Intensity Weights.xlsx", range = cell_rows(10:767), col_names=TRUE)

colnames(intensity)[which(colnames(intensity)=="Pharmacy Intensity Weight")] <- "pharma_wgt"
colnames(intensity)[which(colnames(intensity)=="Supply Intensity Weight")] <- "supply_wgt"

intensity <- intensity[c("MS-DRG","pharma_wgt", "supply_wgt")]

patUse <- merge(patUse, intensity, by="MS-DRG", all.x=TRUE)
summary(patUse$pharma_wgt)
summary(patUse$supply_wgt)

# unmatched DRGs?
test <- subset(patUse, is.na(pharma_wgt))
table(test$`MS-DRG`) # 90% unmatched have DRG 795 - normal newborn (16941/19056)
table(test$`MS-DRG_Desc`)
table(test$DischargeDateMonth)
```

## Create outcome vars

### Cost outcomes
- Main outcome: `Enc - Total Variable Direct Cost`
- TVDC for subcategories:
  - Blood
  - ICU/CCU
  - Implants
  - Laboratory
  - MRI
  - Pharmacy
  - Radiology
- Cost variability: measure cost variability later using the coefficient of variation (only at the aggregate level)

create sum of ICU/CCU + Post Op/Stepdown & Implants + Supplies
```{r}
patUse$ICU_postop_cpi <- patUse$ICU_cpi + patUse$postop_cpi
patUse$implants_supplies_cpi <- patUse$Implants_cpi + patUse$supplies_cpi
```

### LOS outcome
- average `Length of Stay`
- observed / expected LOS using `NYU Vizient Expected LOS`  

create observed / expected LOS using `NYU Vizient Expected LOS`  
```{r}
colnames(patUse)[which(colnames(patUse)=="NYU Vizient Expected LOS")] <- "explos"
table(patUse$explos)
# 34381 have 0 expected LOS

patUse$oelos <- patUse$`Length of Stay` / (patUse$explos)
```

create percent of cases with O/E LOS > 1.5 ("12% down to 9%") (later average the indicators for each month)
```{r}
hist(patUse$oelos)
summary(patUse$oelos)
patUse$oelos_gt15 <- ifelse(patUse$oelos > 1.5, 1, 0)
```

### create counterbalancing health outcome vars

create readmission outcome
```{r}
patUse$readmit <- ifelse(patUse$`NYU Vizient Readmit Flag`=="Y" & !is.na(patUse$`NYU Vizient Readmit Flag`), 1, 0)
summary(patUse$readmit) # 6.3% readmission
```

create mortality outcome
```{r}
patUse$death <- ifelse(patUse$`Discharge Disposition_Desc`=="Expired" | patUse$`Discharge Disposition_Desc`=="EXPIRED", 1, 0)
summary(patUse$death) # 0.67% mortality rate

# combine death + hospice
table(patUse$`Discharge Disposition_Desc`)
patUse$death_hospice <- ifelse(patUse$`Discharge Disposition_Desc`=="Expired" | patUse$`Discharge Disposition_Desc`=="EXPIRED"| patUse$`Discharge Disposition_Desc`=="HOSPICE/HOME" | patUse$`Discharge Disposition_Desc`=="HOSPICE/MEDICAL FACILITY", 1, 0)
summary(patUse$death_hospice)
# 1.7% mortality / hospice rate
```


### sample restriction
```{r}
patUse2 <- patUse
```

drop 3 obs with LOS = 0
```{r}
# rename LOS column name
colnames(patUse2)[which(colnames(patUse2)=="Length of Stay")] <- "los"
patUse2 <- patUse2[patUse2$los > 0 ,]
```

drop where surgical is missing
```{r}
patUse2 <- patUse2[!is.na(patUse2$surgical),]
```

drop where insurance is missing
```{r}
patUse2 <- patUse2[!is.na(patUse2$ins),]
```

drop where mortality indicator is missing
```{r}
patUse2 <- patUse2[!is.na(patUse2$death),]
patUse2 <- patUse2[!is.na(patUse2$death_hospice),]
```

Quick summary stats to check for abnormal values (e.g. negative cost values)
```{r}
nums <- sapply(patUse2, is.numeric)
summary(patUse2[,nums])
```

count and remove discharges with (-) in outcome/age vars, where they shouldn't
```{r}
outcome <- c(costs, "Age", "los")
summary(patUse2[outcome])

# count minus obs
for(i in 1:length(outcome)){
  x <- sum(patUse2[outcome[i]] < 0)
  cat(outcome[i],' has (-) values in ', x, 'obs \n')
}

# remove minus obs
for(i in 1:length(outcome)){
  index <- which(patUse2[,outcome[i]] >= 0)
  patUse2 <- patUse2[index, ]
}
# 658 obs dropped
```

exclude TVDC < $100; cost values can't be too small
```{r}
table(patUse2$tvdc)

patUse2 <- patUse2[patUse2$tvdc >100, ]
summary(patUse2$tvdc)
```


exclude outlier costs = top 0.05%
```{r}
# create z-score of the log cost
# patUse2$logtvdc_cpi <- log(patUse2$tvdc_cpi)
# patUse2$logtvdc_z <- scale(patUse2$logtvdc_cpi, center = TRUE, scale = TRUE)
# summary(patUse2$logtvdc_z)

quantile(patUse2$tvdc_cpi, 0.9995)
quantile(patUse2$tvdc_cpi, 0.9995) / median(patUse2$tvdc_cpi) # bigger than median by a factor of 41

patUse2$big <- patUse2$tvdc_cpi > quantile(patUse2$tvdc_cpi, 0.9995)

n <- sum(patUse2$big)

test <- subset(patUse2, (DischargeDateMonth < "201209" | DischargeDateMonth > "201308") & !(DischargeDateMonth>="201404" & DischargeDateMonth<="201408"))

# scatterplot of CMI vs TVDC
library(ggplot2)
library(dplyr)
ggplot(test, aes(x=CMI, y=tvdc_cpi, colour=factor(big, labels=c("Not outlier","Cost outlier")))) +
  geom_point() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="CPI-adjusted total variable direct cost ($)",
       caption=paste0("I exclude the ", n," outlier observations."))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_cost.png")

ggplot(test[test$big==0,], aes(x=CMI, y=tvdc_cpi)) +
  geom_point() +
  # geom_smooth() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="CPI-adjusted total variable direct cost ($)",
       caption=paste0("I exclude the ", n," outlier observations."))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_cost_nooutlier.png")

ggplot(test[test$big==0,], aes(x=CMI, y=tvdc_cpi, colour=factor(int, labels=c("Pre", "Post")))) +
  # facet_grid(.~factor(`Fiscal Year`)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="Inflation-adjusted total variable direct cost ($)",
       title=("Total variable direct cost vs CMI"))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_cost_byint.png")

patUse2 <- patUse2[patUse2$big==0,]
```

create categorical CMI group vars
```{r}
patUse2$cmiGroup <- cut(patUse2$CMI, breaks=c(quantile(patUse2$CMI, probs = seq(0, 1, by = 0.25))),
      labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)
quantile(patUse2$CMI, probs = seq(0, 1, by = 0.25))
 #     0%     25%     50%     75%    100%
 # 0.1649  0.7569  1.2494  2.0953 26.2466
```

create CMI deciles 
```{r}
patUse2$cmiD <- cut(patUse2$CMI, breaks=c(quantile(patUse2$CMI, probs = seq(0, 1, by = 0.1), na.rm=TRUE)),
  labels=c("0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90","90-100"), include.lowest=TRUE)
table(patUse2$cmiD)
```

merge in elixhauser dummies
```{r}
# export the patient data to get elixhauser variables
v <- c("PATIENT ID", "Enc - Primary ICD10 Diagnosis", "Enc - Primary ICD10 Surgical Procedure", "Enc - Primary ICD9 Diagnosis", "Enc - Primary ICD9 Surgical Procedure", "Discharge Date")
pat_elix_03162018 <- patUse2[v]
write.csv(pat_elix_03162018, file="~/Box Sync/VBM/Data/pat_elix_03162018.csv")

# run crelixhauser.do on server in Stata

elix <- read.csv("~/Box Sync/VBM/Data/elixhauser.csv", header=TRUE)

# merge with discharge-level data
NROW(patUse2)
patUse2 <- merge(patUse2, elix, by.x="PATIENT ID", by.y="PATIENT_ID", all.x=TRUE)
```

drop if discharge date month is missing
```{r}
patUse2 <- patUse2[complete.cases(patUse2[, c("DischargeDateMonth")]),]
```

save as project
```{r}
save(pat, pat2, pat3, patUse, patUse2, dateLookUp, file="~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")

write.csv(patUse2, "~/Box Sync/VBM/Data/patUse2.csv", col.names=TRUE)
```



## Discharge-level analysis
```{r}
load("~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")

```

### Sub-category costs are variable costs only? 
```{r}
str(patUse2)
```


### Define some useful variables
```{r}
costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU_postop", "implants_supplies","Blood", "OR", "inhalation","POT", "routine")

costs_lab <- c("Total variable direct", "Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU/CCU and Post Op/Stepdown", "Implants and Medical/Surgical supplies","Blood", "OR", "Inhalation therapy","Physical/Occupational therapy", "Routine care")

costs_cpi <- paste(costs, "_cpi", sep="")
costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")

for (col in costs_cpi) {
  y <- paste0(col, "_cmi")
  patUse2[,y] <- patUse2[,col] / patUse2$CMI
}

costs <- c("Pharmacy")
costs_cpi <- paste(costs, "_cpi", sep="")
costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
for (col in costs_cpi) {
  y <- paste0(col, "_cmi")
  patUse2[,y] <- patUse2[,col] / patUse2$pharma_wgt
}
```


### Descriptive stats of outcomes
```{r}
patUse3 <- subset(patUse2, DischargeDateMonth < "201209" | DischargeDateMonth > "201308")

costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU_postop", "implants_supplies","Blood", "OR", "inhalation","POT", "routine")
costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")

stargazer(patUse3[costs_cpi_cmi], type = "html",
          title="Descriptive statistics", digits=0,
          out="~/Dropbox/Research/VBM/results/summstats_cost.htm",
          covariate.labels=costs_lab,
          summary.stat = c("n", "mean", "sd", "min", "p25", "median", "p75", "max"))

# Pr(subcategory cost > 0)
summary(patUse3[costs_cpi_cmi])
for (col in costs_cpi_cmi) {
  n <- sum(patUse3[col]==0)
  print(paste0(col, ": ", n, " discharges have 0 costs"))
  x <- paste0(col,"_pos")
  patUse3[,x] <- ifelse(patUse3[,col]==0,0,1)
}

costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU_postop", "implants_supplies","Blood", "OR", "inhalation","POT", "routine")
costs_cpi_cmi_pos <- paste(costs, "_cpi_cmi_pos", sep="")

stargazer(patUse3[costs_cpi_cmi_pos], type = "html",
          title="Descriptive statistics", digits=2,
          out="~/Dropbox/Research/VBM/results/summstats_subcost_pos.htm",
          covariate.labels=costs_lab,
          summary.stat = c("n", "mean"))

# conditional on being positive, summ stats on the sub costs
init <- 0
for (col in costs_cpi_cmi) {
  init <- init + 1
  x <- paste0(col,"_pos")
  patUse4 <- patUse3[patUse3[,x]==1,]
  # print(summary(patUse4[,col]))

  stargazer(patUse4[col], type = "html",
          title="Descriptive statistics", digits=0,
          out=paste0("~/Dropbox/Research/VBM/results/summstats_", col,".htm"),
          covariate.labels=costs_lab[init],
          summary.stat = c("n", "mean", "sd", "min", "p25", "median", "p75", "max"))
}

stargazer(patUse3[c("los", "death")], type = "html",
          title="Descriptive statistics", digits=3,
          out="~/Dropbox/Research/VBM/results/summstats2.htm",
          covariate.labels=c("Length of stay (LOS) (days)", "Indicator for in-hospital death"),
          summary.stat = c("n", "mean", "sd", "min", "p25", "median", "p75", "max"))

table(patUse3$DischargeDateMonth, patUse3$explos==0)

stargazer(patUse3[patUse3$DischargeDateMonth >= "201309" & patUse3$explos > 0, c("oelos", "oelos_gt15")], type = "html",
          title="Descriptive statistics", digits=3,
          out="~/Dropbox/Research/VBM/results/summstats3.htm",
          covariate.labels=c("Observed/expected (O/E) LOS ratio", "Indicator for O/E LOS ratio > 1.5"),
          summary.stat = c("n", "mean", "sd", "min", "p25", "median", "p75", "max"))

test <- patUse3[patUse3$DischargeDateMonth >= "201201", c("readmit")]
stargazer(as.data.frame(test), type = "html",
          title="Descriptive statistics", digits=3,
          out="~/Dropbox/Research/VBM/results/summstats4.htm",
          covariate.labels=c("Indicator for hospital readmission"),
          summary.stat = c("n", "mean", "sd", "min", "p25", "median", "p75", "max"))
```

### stacked bar chart of # discharges by patient type across months
```{r}
# by patient type
test <- ddply(patUse2, .(DischargeDateMonth, pattype), summarise, N=length(pattype))
test$N[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308"] <- NA

# scatterplot
ggplot(test, aes(x=DischargeDateMonth, y=N, colour=pattype, group=pattype)) +
  theme_bw(base_size=15) +
  geom_line() +
  geom_point() +
  labs(x="Month", y="Number of discharges",
       title="Number of medical, surgical patient discharges by month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  # geom_vline(aes(xintercept=36), colour="red") +
  scale_y_continuous(limits = c(0,2000),
                     breaks=seq(0,2000,200)) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))

# bar chart
ggplot(test) +
  theme_bw(base_size=13) +
  geom_bar(mapping=aes(x=DischargeDateMonth, y=N, fill=pattype, group=pattype), stat="identity") +
  labs(x="Month", y="Number of discharges",
       title="Number of Medical and Surgical Patient Discharges by Month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  # geom_vline(aes(xintercept=36), colour="red") +
  scale_y_continuous(limits = c(0,3500),
                     breaks=seq(0,3500,500)) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/num_discharges_bymedsurg.png", width=20, height=10, units="cm")
```

### plot the mean cost & CMI trajectory 

All patients
```{r}
costs <- c("tvdc")
costs_lab <- c("Total variable direct")
costs_cpi <- paste(costs, "_cpi", sep="")
costs_cpi_cmi <- paste(costs_cpi, "_cmi", sep="")

keep <- c(costs, costs_cpi, costs_cpi_cmi, "DischargeDateMonth", "CMI")
test3 <- ddply(patUse2[keep], .(DischargeDateMonth), numcolwise(mean, na.rm=TRUE))

for (col in c(costs, costs_cpi, costs_cpi_cmi, "CMI")) {
  test3[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308", col] <- NA
}

# just drop FY 13
# test3 <- subset(test3, !(DischargeDateMonth >= "201209" & DischargeDateMonth <= "201308"))

# line graphs: build up and show raw, just inflation adjusted, and CMI & inflation adjusted
library(reshape2)
test <- melt(test3[c("tvdc", "tvdc_cpi", "tvdc_cpi_cmi", "DischargeDateMonth")], id.var="DischargeDateMonth")
test$variable <- factor(test$variable, labels = c("Raw", "Inflation-adjusted", "Inflation and DRG weight-adjusted"))

# all 3 costs together
ggplot(test, aes(x=DischargeDateMonth, y=value, colour=variable, group=variable)) +
  theme_bw(base_size=13) +
  geom_line() +
  geom_point() +
  labs(x="Month", y="Mean cost per discharge ($)",
       title="Mean Total Variable Direct Cost per Discharge by Month") +
       # caption="Notes. May 2014 - August 2014 is the transition period.") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  # guides(col = guide_legend(ncol = 1)) +
  geom_vline(aes(xintercept=32), colour="red") +
  # geom_vline(aes(xintercept=21), colour="red") +
  # geom_vline(aes(xintercept=25), colour="red") +
  scale_y_continuous(limits = c(0,12500),
                     breaks=seq(0,12500,2000)) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/tvdc_trajectory_all3.pdf", width=20, height=10, units="cm")

#only the inflation & DRG weight adjusted TVDC
ggplot(test[test$variable=="Inflation and DRG weight-adjusted",], aes(x=DischargeDateMonth, y=value, group=variable)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Mean cost per discharge ($)",
       title="Mean Total Variable Direct Cost per Discharge by Month",
       subtitle="All patients") +
       # caption="Notes. May 2014 - August 2014 is the transition period.") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  scale_y_continuous(limits = c(0,6100),
                     breaks=seq(0,6100,1000)) +
    scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/tvdc_cpi_cmi_trajectory.pdf", width=20, height=10, units="cm")
```

aggregate mean CMI by month
```{r}
ggplot(test3, aes(x=DischargeDateMonth, y=CMI, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Mean DRG weights",
       title="Mean DRG Weights by Month",
       subtitle="All Patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,2) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/CMI_trajectory.pdf", width=20, height=10, units="cm")
```


plot separately by patient type
```{r}
keep <- c(costs, costs_cpi, costs_cpi_cmi, "DischargeDateMonth", "CMI", "pattype")
test3 <- ddply(patUse2[keep], .(DischargeDateMonth, pattype), numcolwise(mean, na.rm=TRUE))

for (col in c(costs, costs_cpi, costs_cpi_cmi, "CMI")) {
  test3[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308", col] <- NA
}

# line graphs: build up and show raw, just inflation adjusted, and CMI & inflation adjusted
library(reshape2)
test3 <- data.table(test3)
test <- melt(test3, id.vars=c("DischargeDateMonth","pattype"), measure.vars=c("tvdc", "tvdc_cpi", "tvdc_cpi_cmi"))
test$variable <- factor(test$variable, labels = c("Raw", "Inflation-adjusted", "Inflation and DRG weight-adjusted"))

ggplot(test[test$variable=="Inflation and DRG weight-adjusted",], aes(x=DischargeDateMonth, y=value, group=pattype, colour=pattype)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Mean cost per discharge ($)",
       title="Mean Total Variable Direct Cost per Discharge by Month",
       subtitle="By Medical and Surgical Patients") +
       # caption="Notes. May 2014 - August 2014 is the transition period.") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  scale_y_continuous(limits = c(0,6100),
                     breaks=seq(0,6100,1000)) +
    scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/trajectory_tvdc_cpi_cmi_pattype.pdf", width=20, height=10, units="cm")
```

aggregate mean CMI by month
```{r}
ggplot(test3, aes(x=DischargeDateMonth, y=CMI, group=pattype, colour=pattype)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Mean DRG weights",
       title="Mean DRG Weights by Month",
       subtitle="By Medical and Surgical Patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,3.5) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/CMI_trajectory_bypattype.pdf", width=20, height=10, units="cm")
```


### plot the CMI trajectory


### Elixhauser comorbidity

Elixhauser comorbidity index trajectory
```{r}
patUse3 <- patUse2

# all patients
keep <- c("CMI", "elixsum", "tvdc_cpi", "DischargeDateMonth", "int")
test3 <- ddply(patUse3[keep], .(DischargeDateMonth), numcolwise(mean, na.rm=TRUE))

for (col in c("CMI", "elixsum", "tvdc_cpi")) {
  test3[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308", col] <- NA
}

ggplot(test3, aes(x=DischargeDateMonth, y=elixsum, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Elixhauser comorbidity index",
       title="Mean Elixhauser Comorbidity Index by Month",
       subtitle="All Patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,0.5) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/elixsum_trajectory.pdf", width=20, height=10, units="cm")

# by patient type
keep <- c("CMI", "elixsum", "tvdc_cpi", "DischargeDateMonth", "pattype")
test3 <- ddply(patUse3[keep], .(DischargeDateMonth, pattype), numcolwise(mean, na.rm=TRUE))

for (col in c("CMI", "elixsum", "tvdc_cpi")) {
  test3[test3$DischargeDateMonth >= "201209" & test3$DischargeDateMonth <= "201308", col] <- NA
}

ggplot(test3, aes(x=DischargeDateMonth, y=elixsum, group=pattype, colour=pattype)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Elixhauser comorbidity index",
       title="Mean Elixhauser Comorbidity Index by Month",
       subtitle="By Medical and Surgical Patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,0.5) +
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))

ggsave("~/Dropbox/Research/VBM/results/elixsum_trajectory_bypattype.pdf", width=20, height=10, units="cm")

```

Elixhauser comorbidity index vs CMI - skip
```{r}
patUse3 <- subset(patUse2, DischargeDateMonth < "201209" | DischargeDateMonth > "201308")

ggplot(patUse3, aes(x=factor(elixsum), y=CMI)) +
  # facet_grid(.~factor(`Fiscal Year`)) +
  geom_boxplot() +
  geom_smooth() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="DRG weights",
       x="Elixhauser comorbidity index",
       title=("DRG weights vs Elixhauser comorbidity index"))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_elixhauser.png")
```


### CMI vs Cost - skip
```{r}
patUse3 <- subset(patUse2, (DischargeDateD < "2012-09-01" | DischargeDateD > "2013-08-31") )

ggplot(patUse3, aes(x=CMI, y=tvdc_cpi, colour=factor(int, labels=c("Pre", "Post")))) +
  # facet_grid(.~factor(`Fiscal Year`)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="Inflation-adjusted total variable direct cost ($)",
       title=("Total variable direct cost vs CMI"))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_cost_byint.png")

ggplot(patUse3, aes(x=cmiGroup, y=tvdc_cpi)) +
  # facet_grid(.~factor(`Fiscal Year`)) +
  geom_boxplot() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="Inflation-adjusted total variable direct cost ($)",
       title=("Total variable direct cost vs CMI quartiles"))
ggsave("~/Dropbox/Research/VBM/results/CMI_vs_cost_byCMIq.png")

patUse3$CMI2 <- (patUse3$CMI)^2

ggplot(patUse3, aes(x=CMI2, y=tvdc_cpi, colour=factor(int, labels=c("Pre", "Post")))) +
  # facet_grid(.~factor(`Fiscal Year`)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(y="Inflation-adjusted total variable direct cost ($)",
       title=("Total variable direct cost vs CMI squared"))
ggsave("~/Dropbox/Research/VBM/results/CMIsq_vs_cost_byint.png")

patUse3$tvdc_cpi_cmi <- patUse3$tvdc_cpi / patUse3$CMI
summary(patUse3$tvdc_cpi_cmi)
summary(patUse3$cmiGroup)

ggplot(patUse3 ) +
  geom_density(aes(tvdc_cpi), colour="red") +
  geom_density(aes(tvdc_cpi_cmi), colour="blue")

ggplot(patUse3, aes(tvdc_cpi_cmi)) +
  geom_density()

```





### Regression analysis

just before running regressions, suppress to NA the discharge dates 9/2012 - 8/2013
```{r}
load("~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")

# patUse3 <- subset(patUse2, DischargeDateMonth < "201209" | DischargeDateMonth > "201308")

# using Sept 2014 as the first post month
patUse3 <- subset(patUse2, DischargeDateMonth < "201209" |(DischargeDateMonth > "201308" & DischargeDateMonth < "201405") | DischargeDateMonth >= "201409")
patUse3$int[patUse3$DischargeDateMonth < "201409"] <- 0
patUse3$tpostInt[patUse3$DischargeDateMonth < "201409"] <- 0
patUse3$tpostInt[patUse3$DischargeDateMonth >= "201409"] <- patUse3$tpostInt[patUse3$DischargeDateMonth >= "201409"] -4

test <- ddply(patUse3, .(DischargeDateMonth), summarise, int = mean(int), tpostInt = mean(tpostInt), t = mean(t))

```

#### Main outcome: Total variable direct cost

```{r}
# elix <- paste0("ynel", 1:31)
# elix <- paste(elix, collapse= " + ")

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

costs_cpi <- ("tvdc_cpi")
costs_lab <- ("Inflation-adjusted total variable direct costs (%)")

# costs_cpi_cmi <- ("tvdc_cpi_cmi")
# patUse3$tvdc_cpi_cmi <- patUse3$tvdc_cpi / patUse3$CMI
# summary(patUse3$tvdc_cpi_cmi)

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse3, ~ site, report = T)
  # result2 <- result
  for (i in 1:n) {
    print(summary(result[[i]]))
  }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,
            # keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
# could report SEs for exponentiated coefficients using the Delta method (see: https://stats.idre.ucla.edu/r/faq/how-can-i-estimate-the-standard-error-of-transformed-regression-parameters-in-r-using-the-delta-method/)

# compute pairs cluster bootstrapped p-values
# library(clusterSEs)
# clust.bs.p <- cluster.bs.glm(its.lm1, patUse4, ~ DischargeDateMonth, report = T)
```

medical & surgical patients separately
```{r}
# for rehab, they have all elixhauser comorbidity indicators = 0 except 1 pat

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

# medical patients
patUse4 <- subset(patUse3, surgical=="0")

costs_cpi <- ("tvdc_cpi")
costs_lab <- ("Inflation-adjusted total variable direct costs (%)")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)
  # result <- lapply(models, FUN = function(x) {
  #   lm(formula = x, data=patUse4, na.action="na.omit")
  # })
  
  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse3, ~ site, report = T)
  # result2 <- result
  for (i in 1:n) {
    print(summary(result[[i]]))
  }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "medical.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}


# surgical patients
comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")


patUse4 <- subset(patUse3, surgical=="1")

# some variables have only 1 level
for (i in 1:31) {
  x <- paste0("ynel",i)
  print(x)
  print(summary(patUse4[,x]))
}
# ynel25 have only 0 one level (absent) (See https://stackoverflow.com/questions/18171246/error-in-contrasts-when-defining-a-linear-model-in-r)
summary(patUse4$ynel25)

costs_cpi <- ("tvdc_cpi")
costs_lab <- ("Inflation-adjusted total variable direct costs (%)")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse3, ~ site, report = T)
  # result2 <- result
  for (i in 1:n) {
    print(summary(result[[i]]))
  }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "surgical.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```


for education, compare with model with only intervention dummy - skip
```{r}
costs_cpi <- ("tvdc_cpi")
costs_lab <- ("Inflation-adjusted total costs (%)")

pct_chng <- function(x) 100*(exp(x)-1)


  lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)
glm(formula = x,
              data = patUse3,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))

init <- 0
n <- length(comb)
for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse3, ~ site, report = T)
  # result2 <- result
  for (i in 1:n) {
    print(summary(result[[i]]))
  }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_compare.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```


#### subcategory costs
```{r}
# costs <- c("tvdc", "Pharmacy", "Laboratory", "Radiology",  "MRI", "Implants","Blood", "ICU")
costs <- c("Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU_postop", "implants_supplies","Blood", "OR", "inhalation","POT", "routine")

costs_lab <- c("Pharmacy", "Laboratory", "Radiology",  "MRI", "ICU/CCU and Post Op/Stepdown", "Implants and Medical/Surgical supplies","Blood", "OR", "Inhalation therapy","Physical/Occupational therapy", "Routine care")

costs_cpi <- paste(costs, "_cpi", sep="")

# for subcategory costs, bunching at zeros -> use two-part gamma models
summary(patUse3[costs_cpi])
for (col in costs_cpi) {
  n <- sum(patUse3[col]==0)
  print(paste0(col, ": ", n, " discharges have 0 costs"))
  x <- paste0(col,"_pos")
  patUse3[,x] <- ifelse(patUse3[,col]==0,0,1)
}
```

for subcategory costs : "Laboratory", "Radiology",  "MRI", "ICU_postop"
```{r}
con <- file("~/Dropbox/Research/VBM/results/test.log")
sink(con, append=TRUE)
sink(con, append=TRUE, type="message")

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")


costs <- c("POT", "routine")
  # c("Laboratory", "Radiology",  "MRI", "ICU_postop", "OR", 
costs_lab <- c( "Physical/Occupational therapy", "Routine care")
  # c("Laboratory", "Radiology",  "MRI", "ICU/CCU and Post Op/Stepdown", "OR",
               
costs_cpi <- paste(costs, "_cpi", sep="")

# costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
# for (col in costs_cpi) {
#   x <- paste0(col,"_cmi")
#   print(x)
#   patUse3[,x] <- patUse3[,col] / patUse3$CMI
# }

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  # Logit on Pr(cost > 0)
  y <- paste0(col, "_pos")
  models <- lapply(paste0( rep(y,n),"~", comb), formula)

  # patUse3$CMIQ <- cut(patUse3$CMI, breaks=c(quantile(patUse3$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                      # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  logitresult <- lapply(models, FUN = function(x) {
      glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })

  # Gamma on costs using (+) costs data  
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  patUse4 <- patUse3[patUse3[,col] > 0, ]

  # patUse4$CMIQ <- cut(patUse4$CMI, breaks=c(quantile(patUse4$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                         # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })

  stargazer(logitresult, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_pos.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  for (i in 1:n) {
    print(col)
    print(summary(logitresult[[i]]))
    print(summary(result[[i]]))
  }
}
```

subcategory costs: "Blood"
```{r}
comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")


# some variables have only 1 level
patUse4 <- subset(patUse3, Blood_cpi > 0)
for (i in 1:31) {
  x <- paste0("ynel",i)
  print(x)
  print(summary(patUse4[,x]))
}
# ynel25 have only 0 one level (absent) (See https://stackoverflow.com/questions/18171246/error-in-contrasts-when-defining-a-linear-model-in-r)
summary(patUse4$ynel29)


costs <- c("Blood")
costs_lab <- c("Blood")
costs_cpi <- paste(costs, "_cpi", sep="")

# costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
# for (col in costs_cpi) {
#   x <- paste0(col,"_cmi")
#   print(x)
#   patUse3[,x] <- patUse3[,col] / patUse3$CMI
# }

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  # Logit on Pr(cost > 0)
  y <- paste0(col, "_pos")
  models <- lapply(paste0( rep(y,n),"~", comb), formula)

  # patUse3$CMIQ <- cut(patUse3$CMI, breaks=c(quantile(patUse3$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                      # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  logitresult <- lapply(models, FUN = function(x) {
      glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })

  # Gamma on costs using (+) costs data  
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  patUse4 <- patUse3[patUse3[,col] > 0, ]

  # patUse4$CMIQ <- cut(patUse4$CMI, breaks=c(quantile(patUse4$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                         # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })

  stargazer(logitresult, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_pos.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  for (i in 1:n) {
    print(col)
    print(summary(logitresult[[i]]))
    print(summary(result[[i]]))
  }
}
```

subcategory costs: "inhalation"
```{r}
comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

# some variables have only 1 level
patUse4 <- subset(patUse3, inhalation_cpi > 0)
for (i in 1:31) {
  x <- paste0("ynel",i)
  print(x)
  print(summary(patUse4[,x]))
}
# ynel25 have only 0 one level (absent) (See https://stackoverflow.com/questions/18171246/error-in-contrasts-when-defining-a-linear-model-in-r)
summary(patUse4$ynel25)


costs <- c("inhalation")
costs_lab <- c("Inhalation therapy")
costs_cpi <- paste(costs, "_cpi", sep="")

# costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
# for (col in costs_cpi) {
#   x <- paste0(col,"_cmi")
#   print(x)
#   patUse3[,x] <- patUse3[,col] / patUse3$CMI
# }

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  # Logit on Pr(cost > 0)
  y <- paste0(col, "_pos")
  models <- lapply(paste0( rep(y,n),"~", comb), formula)

  # patUse3$CMIQ <- cut(patUse3$CMI, breaks=c(quantile(patUse3$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                      # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  logitresult <- lapply(models, FUN = function(x) {
      glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })

  # Gamma on costs using (+) costs data  
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  patUse4 <- patUse3[patUse3[,col] > 0, ]

  # patUse4$CMIQ <- cut(patUse4$CMI, breaks=c(quantile(patUse4$CMI, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
                         # labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1)))
  })

  stargazer(logitresult, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_pos.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  for (i in 1:n) {
    print(col)
    print(summary(logitresult[[i]]))
    print(summary(result[[i]]))
  }
}
```


pharmacy costs
```{r}
costs <- c("Pharmacy")
costs_lab <- c("Pharmacy")
costs_cpi <- paste(costs, "_cpi", sep="")

# costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
# for (col in costs_cpi) {
#   x <- paste0(col,"_cmi")
#   print(x)
#   patUse3[,x] <- patUse3[,col] / patUse3$pharma_wgt
# }

patUse3$pharma_wgtD <- cut(patUse3$pharma_wgt, breaks=c(unique(quantile(patUse3$pharma_wgt, probs = seq(0, 1, by = 0.1), na.rm=TRUE))) ,labels=c("0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-100"), include.lowest=TRUE)
table(patUse3$pharma_wgtD)

patUse4 <- subset(patUse3, !is.na(pharma_wgtD))

# ggplot(patUse4, aes(x=pharma_wgt, y=Pharmacy_cpi)) +
#   geom_point()

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + pharma_wgtD",
          "t + int + tpostInt + season + pharma_wgtD  + surgical",
          "t + int + tpostInt + season + pharma_wgtD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + pharma_wgtD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + pharma_wgtD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  # Logit on Pr(cost > 0)
  y <- paste0(col, "_pos")
  models <- lapply(paste0( rep(y,n),"~", comb), formula)

#   patUse4$pharmaQ <- cut(patUse4$pharma_wgt, breaks=c(quantile(patUse4$pharma_wgt, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
#                          labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  logitresult <- lapply(models, FUN = function(x) {
      glm(formula = x,
              data = patUse4,
              family=binomial(link="logit"),
              na.action="na.omit")
  })

  # Gamma on costs using (+) costs data
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  patUse4 <- patUse4[patUse4[,col] > 0, ]

  # patUse4$pharmaQ <- cut(patUse4$pharma_wgt, breaks=c(quantile(patUse4$pharma_wgt, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
  #     labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse4,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,k-1))
        )
  })

  stargazer(logitresult, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_pos.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
  for (i in 1:n) {
    print(col)
    print(summary(logitresult[[i]]))
    print(summary(result[[i]]))
  }
}
```

supplies costs
```{r}
costs <- c("implants_supplies")
costs_lab <- c("Implants and Medical/Surgical supplies")
costs_cpi <- paste(costs, "_cpi", sep="")

# costs_cpi_cmi <- paste(costs, "_cpi_cmi", sep="")
# for (col in costs_cpi) {
#   x <- paste0(col,"_cmi")
#   print(x)
#   patUse3[,x] <- patUse3[,col] / patUse3$CMI
# }
# patUse3$implants_supplies_cpi_cmi[patUse3$CMI==0] <- NA

patUse3$supply_wgtD <- cut(patUse3$supply_wgt, breaks=c(quantile(patUse3$supply_wgt, probs = seq(0, 1, by = 0.1), na.rm=TRUE)),
  labels=c("0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90","90-100"), include.lowest=TRUE)
table(patUse3$supply_wgtD)

# patUse4 <- subset(patUse3, Pharmacy_cpi > 0 & !is.na(pharma_wgt))
# ggplot(patUse4, aes(x=pharma_wgt, y=Pharmacy_cpi)) +
#   geom_point()

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + supply_wgtD",
          "t + int + tpostInt + season + supply_wgtD  + surgical",
          "t + int + tpostInt + season + supply_wgtD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + supply_wgtD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + supply_wgtD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

pct_chng <- function(x) 100*(exp(x)-1)

patUse4 <- subset(patUse3, !is.na(supply_wgtD))

init <- 0
n <- length(comb)

for (col in costs_cpi) {
  init <- init + 1
  depv <- c(costs_lab[init])

  # Logit on Pr(cost > 0)
  y <- paste0(col, "_pos")
  models <- lapply(paste0( rep(y,n),"~", comb), formula)

  # patUse3$supplyQ <- cut(patUse3$supply_wgt, breaks=c(quantile(patUse3$supply_wgt, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
  #                        labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  logitresult <- lapply(models, FUN = function(x) {
      glm(formula = x,
              data = patUse4,
              family=binomial(link="logit"),
              na.action="na.omit")
  })

  # Gamma on costs using (+) costs data
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  patUse4 <- patUse4[patUse4[,col] > 0, ]

  # patUse4$supplyQ <- cut(patUse4$supply_wgt, breaks=c(quantile(patUse4$supply_wgt, probs = seq(0, 1, by = 0.25), na.rm=TRUE)),
  #                        labels=c("0-25","25-50","50-75","75-100"), include.lowest=TRUE)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse4, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
        data = patUse4,
        family=Gamma(link="log"),
        na.action="na.omit",
        start=c(20, rep(0,k-1)))
  })

  stargazer(logitresult, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, "_pos.htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, "2.htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))

   for (i in 1:n) {
    print(col)
    print(summary(logitresult[[i]]))
    print(summary(result[[i]]))
  }
}

```


#### LOS outcomes: LOS, O/E Ratio, dummy for (O/E Ratio > 1.5)

plot the LOS trajectory over time
```{r}
keep <- c("DischargeDateMonth", "los", "oelos", "oelos_gt15")

patUse3 <- patUse2

# suppress to NA if expected LOS = 0
for (col in c("oelos", "oelos_gt15")) {
  patUse3[patUse3$explos==0, col] <- NA
}

test <- ddply(patUse3[keep], .(DischargeDateMonth), numcolwise(mean, na.rm=TRUE))

for (col in keep[2:4]) {
  test[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308", col] <- NA
}

# LOS trajectory
ggplot(test, aes(x=DischargeDateMonth, y=los, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Length of stay (days)",
       title="Mean length of stay by month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,5) + 
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/los_trajectory.pdf", width=20, height=10, units="cm")

# O/E ratio trajectory
ggplot(test[test$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=oelos, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Observed/expected LOS ratio",
       title="Mean observed/expected LOS ratio by month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=9), colour="red") +
  ylim(0,1.5) + 
  scale_x_discrete(breaks=c("201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/oelos_trajectory.pdf", width=20, height=10, units="cm")

# O/E ratio > 1.5 trajectory
ggplot(test[test$DischargeDateMonth >= "201309",], aes(x=DischargeDateMonth, y=oelos_gt15, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Proportion of discharges",
       title="Proportion of discharges having observed/expected LOS ratio > 1.5") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=9), colour="red") +
  ylim(0,0.5) +
  scale_x_discrete(breaks=c("201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/oelos_gt15_trajectory.pdf", width=20, height=10, units="cm")
```

LOS
```{r}
los_y <- c("los")
los_y_lab <- c("Length of stay (days)")

patUse3 <- subset(patUse2, (DischargeDateD < "2012-09-01" | DischargeDateD > "2013-08-31"))

# exclude O/E ratio > 99th pctile
summary(patUse3$los)
quantile(patUse3$los, c(.99))

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in los_y) {
  init <- init + 1
  depv <- c(los_y_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(5, rep(0,k-1)))
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse2, ~ site, report = T)
  # result2 <- result
  # for (i in 1:n) {
  #   result2[[i]]$coefficients2 <- 100*(exp(result[[i]]$coefficients)-1)
  #   # result2[[i]]$CI.vector <- exp(confint(result[[i]])) - 1
  # }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```

O/E ratio
```{r}
los_y <- c("oelos")
los_y_lab <- c("Observed/Expected (O/E) LOS ratio")

# are expected LOS = 0 for certain months? yes - expected LOS is always 0 before 2012/09 & in 2017/12
table(patUse2$DischargeDateMonth, patUse2$explos==0)

# should use pre-period 2013/09 - 2014/04 only when outcome = O/E ratio
patUse3 <- subset(patUse2, explos > 0 & DischargeDateMonth >= "201309") # drop obs before 2013/09 & 2017/12
table(patUse3$DischargeDateMonth)

# just recode the linear trend term so that it starts from 1 in 2013/09
summary(patUse3$t[patUse3$DischargeDateMonth == "201309"]) # 25
patUse3$t <- patUse3$t-24
table(patUse3$DischargeDateMonth, patUse3$t)
table(patUse3$DischargeDateMonth, patUse3$int)
table(patUse3$DischargeDateMonth, patUse3$tpostInt)

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")


pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in los_y) {
  init <- init + 1
  depv <- c(los_y_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=Gamma(link="log"),
              na.action="na.omit")
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse2, ~ site, report = T)
  # result2 <- result
  # for (i in 1:n) {
  #   result2[[i]]$coefficients2 <- 100*(exp(result[[i]]$coefficients)-1)
  #   # result2[[i]]$CI.vector <- exp(confint(result[[i]])) - 1
  # }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```


dummy for O/E LOS ratio
```{r}
los_y <- c("oelos_gt15")
los_y_lab <- c("Indicator for having O/E LOS ratio > 1.5")

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in los_y) {
  init <- init + 1
  depv <- c(los_y_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse2, ~ site, report = T)
  # result2 <- result
  # for (i in 1:n) {
  #   result2[[i]]$coefficients2 <- 100*(exp(result[[i]]$coefficients)-1)
  #   # result2[[i]]$CI.vector <- exp(confint(result[[i]])) - 1
  # }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```

#### counterbalancing health outcome vars: readmit, death

plot readmission & mortality trajectory
```{r}
patUse3 <- patUse2

keep <- c("DischargeDateMonth", "readmit", "death")
test <- ddply(patUse3[keep], .(DischargeDateMonth), numcolwise(mean, na.rm=TRUE))
# test <- merge(test, dateLookUp, by.x ="DischargeDateMonth", by.y = "dayMonth", all.y=TRUE)

for (col in keep[2:length(keep)]) {
  test[test$DischargeDateMonth >= "201209" & test$DischargeDateMonth <= "201308", col] <- NA
}
test$readmit <- 100*test$readmit
test$death <- 100*test$death

# since readmission is not reported before 2012/01, exclude those months
ggplot(test[test$DischargeDateMonth >= "201201",], aes(x=DischargeDateMonth, y=readmit, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="Readmission rate (%)",
       title="Readmission Rate (%) by Month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=28), colour="red") +
  ylim(0,15) +
  scale_x_discrete(breaks=c("201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/readmit_trajectory.pdf", width=20, height=10, units="cm")

# mortality
ggplot(test, aes(x=DischargeDateMonth, y=death, group=1)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=13) +
  labs(x="Month", y="In-hospital mortality rate (%)",
       title="In-Hospital Mortality Rate (%) by Month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=32), colour="red") +
  ylim(0,3) + 
  scale_x_discrete(breaks=c("201109","201201","201205","201209","201301","201305","201309","201401","201405","201409","201501","201505","201509","201601","201605","201609","201701","201705","201709"))
ggsave("~/Dropbox/Research/VBM/results/death_trajectory.pdf", width=20, height=10, units="cm")
```

```{r}
health_y <- c("readmit")
health_y_lab <- c("Indicator for readmission")

# are expected LOS = 0 for certain months? yes - expected LOS is always 0 before 2012/09
table(patUse2$DischargeDateMonth, patUse2$readmit==1) # readmission=0 always before 2012/01
table(patUse2$DischargeDateMonth, patUse2$death==1) # readmission=0 always before 2012/01

# should use pre-period 2013/09 - 2014/04 only when outcome = O/E ratio
patUse3 <- patUse2[patUse2$DischargeDateMonth >= "201201" ,]

patUse3 <- subset(patUse3, DischargeDateD < "2012-09-01" | DischargeDateD > "2013-08-31" )

table(patUse3$DischargeDateMonth)

# just recode the linear trend term so that it starts from 1 in 201201
summary(patUse3$t[patUse3$DischargeDateMonth == "201201"]) # 5
patUse3$t <- patUse3$t-5 + 1
table(patUse3$DischargeDateMonth, patUse3$t)

# # standardize the values
# table(patUse4$`Discharge Service_Desc`)
# patUse4$disch_dpt <- patUse4$`Discharge Service_Desc`
# patUse4$disch_dpt < as.character(patUse4$disch_dpt)
# patUse4$disch_dpt[patUse4$disch_dpt=="Unk"] <- NA
# patUse4$disch_dpt[patUse4$disch_dpt=="RADIOLOGY"] <- "Radiology"
# patUse4$disch_dpt[patUse4$disch_dpt=="OPHTHALMOLOGY"] <- "Ophthalmology"
# patUse4$disch_dpt[patUse4$disch_dpt=="NEUROLOGY"] <- "Neurology"
# patUse4$disch_dpt[patUse4$disch_dpt=="Gynecology"] <- "OB/GYN-GYN"
# patUse4$disch_dpt[patUse4$disch_dpt=="PSYCHIATRY"] <- "Psychiatry"
# patUse4$disch_dpt[patUse4$disch_dpt=="OTOLARYNGOLOGY"] <- "Otolaryngology"
# disch_dpt_df <- count(patUse4, 'disch_dpt')
# write.csv(disch_dpt_df, file="~/Dropbox/Research/VBM/results/disch_dpt.csv")

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in health_y) {
  init <- init + 1
  depv <- c(health_y_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse2, ~ site, report = T)
  # result2 <- result
  # for (i in 1:n) {
  #   result2[[i]]$coefficients2 <- 100*(exp(result[[i]]$coefficients)-1)
  #   # result2[[i]]$CI.vector <- exp(confint(result[[i]])) - 1
  # }
  # for (i in 1:n) {
  #   result[[i]]$coefficients[,4] <- cluster.bs.glm(result[[i]], col, ~ area, report = T)
  # }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}

# mixed effects logistic regression with random effects from discharge service dept
```

death
```{r}
health_y <- c("death")
health_y_lab <- c("Indicator for in-hospital death")

comb <- c("t + int + tpostInt + season",
          "t + int + tpostInt + season + cmiD",
          "t + int + tpostInt + season + cmiD  + surgical",
          "t + int + tpostInt + season + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31",
          "t + int + tpostInt + season + cmiD  + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup",
          "t + int + tpostInt + season  + cmiD + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins")

patUse3 <- subset(patUse2, DischargeDateD < "2012-09-01" | DischargeDateD > "2013-08-31" )


pct_chng <- function(x) 100*(exp(x)-1)

init <- 0
n <- length(comb)
for (col in health_y) {
  init <- init + 1
  depv <- c(health_y_lab[init])

  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    k <- length(lmcoef)

    glm(formula = x,
              data = patUse3,
              family=binomial(link="logit"),
              na.action="na.omit")
  })
  # clust.bs.p <- cluster.bs.glm(result[[i]], patUse2, ~ site, report = T)
  # result2 <- result
  # for (i in 1:n) {
  #   result2[[i]]$coefficients2 <- 100*(exp(result[[i]]$coefficients)-1)
  #   # result2[[i]]$CI.vector <- exp(confint(result[[i]])) - 1
  # }
  # for (i in 1:n) {
  #   result[[i]]$coefficients[,4] <- cluster.bs.glm(result[[i]], col, ~ area, report = T)
  # }

  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```

### stop sinking file (i.e. saving console output to log file)
```{r}
sink(file=NULL)
```


## SPC of inflation & CMI adjsuted total costs
```{r}
library(qcc)

patUse2$tvdc_cpi_cmi <- patUse2$tvdc_cpi / patUse2$CMI

patUse3 <- subset(patUse2, (DischargeDateMonth < "201209" | DischargeDateMonth > "201308"))

pdf("~/Dropbox/Research/VBM/results/qcc_tvdc_cpi.pdf")
my.xmr.x <- qcc(data =  qcc.groups(patUse3$tvdc_cpi[patUse3$int==0],
                                   patUse3$t[patUse3$int==0]),
                 newdata= qcc.groups(patUse3$tvdc_cpi[patUse3$int==1],
                                     patUse3$t[patUse3$int==1]),
                title="Statistical Process Control model: \nMean inflation-adjusted total direct variable cost per discharge by month",
                # labels= c(seq.Date(from=as.Date("2010/09/01"),to = as.Date("2012/08/31"), by="month" ), seq.Date(from=as.Date("2013/09/01"),to = as.Date("2017/05/31"), by="month" )),
                type = "xbar",
                ylim=c(0,14000),
                ylab="Total direct variable cost ($)")
lines(rep(21,10), seq(0,15000,length.out=10), col="blue")
dev.off()

pdf("~/Dropbox/Research/VBM/results/qcc_tvdc_cpi_cmi.pdf")
my.xmr.x <- qcc(data =  qcc.groups(patUse3$tvdc_cpi_cmi[patUse3$int==0],
                                   patUse3$t[patUse3$int==0]),
                 newdata= qcc.groups(patUse3$tvdc_cpi_cmi[patUse3$int==1],
                                     patUse3$t[patUse3$int==1]),
                title="Statistical Process Control model: \nMean inflation and DRG weight-adjusted total direct variable cost per discharge",
                # labels= c(seq.Date(from=as.Date("2010/09/01"),to = as.Date("2012/08/31"), by="month" ), seq.Date(from=as.Date("2013/09/01"),to = as.Date("2017/05/31"), by="month" )),
                type = "xbar",
                ylim=c(0,8000),
                ylab="Total direct variable cost per discharge ($)")
lines(rep(21,10), seq(0,15000,length.out=10), col="blue")
dev.off()

# my.xmr.x$center
# my.xmr.x$limits[,1]
# my.xmr.x$limits[,2]
```


## Aggregate analysis at the monthly/weekly level
The following analysis is at the aggregate level using monthly data.  

export patient discharge level data to get coefficients on month dummies while controlling for patient vars
```{r}
load("~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")

patUse3 <- subset(patUse2, (DischargeDateMonth < "201209" | DischargeDateMonth > "201308"))
write.csv(patUse3, file="~/Box Sync/VBM/Data/patUse3.csv")

```



### Create the monthly time-series data of risk-adjusted TVDC 

import risk-adjusted total costs, which are month fixed effect coefficientss estimated with patient characteristics controls using the discharge-level data in Stata (VBManalysis.do)

```{r}

# import risk-adjusted total costs, which are month fixed effect coefficientss estimated with patient characteristics controls using the discharge-level data in Stata (VBManalysis.do)
racost <- read.csv("~/Dropbox/Research/VBM/data/ra_tvdc_cpi_monthly.csv")
library(zoo)
racost$tt <- as.yearmon(paste(racost$yr, racost$month, sep = "-"))
racost$DischargeDateMonth <- format(racost$tt, "%Y%m")
colnames(racost)[which(colnames(racost)=="ra_month")] <- "tvdc_cpi"
colnames(racost)[which(colnames(racost)=="ra_month_surg0")] <- "tvdc_cpi_surg0"
colnames(racost)[which(colnames(racost)=="ra_month_surg1")] <- "tvdc_cpi_surg1"
racost <- racost[c("tvdc_cpi", "tvdc_cpi_surg1", "tvdc_cpi_surg0", "DischargeDateMonth")]

# add in FY 13
allDates <- as.data.frame(sort(unique(patUse2$DischargeDateMonth)))
names(allDates) <- "DischargeDateMonth"

lastpremonthidx <- 31
nmonths <- length(allDates$DischargeDateMonth)

allDates$int <- c(rep(0, lastpremonthidx), rep(1, nmonths-lastpremonthidx))
allDates$t <- 1:nmonths

# aggregate data
racost <- merge(racost, allDates, by="DischargeDateMonth", all.y=TRUE)

# time-series data
aggData.ts <- ts(racost, start=c(2011,9), end=c(2017,5), frequency = 12)
summary(aggData.ts)

plot(aggData.ts[,c("tvdc_cpi")], ylim=c(0,10000) ) # plot avg total variable direct cost
```


### ITS model using the patient risk adjusted TVDC

Work out AR / MA order of time series
```{r}
costs_cpi <- c("tvdc_cpi","tvdc_cpi_surg1", "tvdc_cpi_surg0")
for (col in costs_cpi) {
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/acf_",col,".pdf", sep=""))
  par(mfrow=c(2,1))
  acf(aggData.ts[,col], na.action = na.pass )
  pacf(aggData.ts[,col], na.action = na.pass )
  dev.off()
}
# PACF order= 1 for all pats; =3 for surg, =5 for med pats
```

Define a function to estimate an ITS model
```{r}
f_mav <- function(x,n=5){stats::filter(x,rep(1/n,n), sides=2)}

f_splitPlot2 <- function ( plotTimeSeries, switchObservation, frequency, firstMonth , lastPreMonth ,
                           firstPostMonth ,  lastMonth ,
                           xlab="Date", ylab="Count", main=NULL, sub=NULL, arOrder,
                           ylim=NULL, r) {
  ## Plot outline with shading
  require(forecast)
  library(nlme)
  base <-  switchObservation # row number for the first post period

  ## Intrupted time series model
  time <- 1:NROW(plotTimeSeries)
  intervention <- c(rep(0,base-1),rep(1,NROW(plotTimeSeries)-base+1))
  timeAfterIntervention  <- c(rep(0,base-1),1:((NROW(plotTimeSeries)-base+1)) )
  its.lm <- gls(plotTimeSeries ~ time + intervention +  timeAfterIntervention,
                correlation=corARMA(p=arOrder), method="ML",
                na.action="na.omit")    
  cf1 <- summary(its.lm)
  cf <- cf1$tTable
  print(cf)

  CI <- confint(its.lm)
  print(CI)

  sub <- paste("Overall trend = ", round(cf[2,1],r) , " (", round(CI[2,1],r), ", ", round(CI[2,2],r), "), p=", round(cf[2,4],r), "\n",
               "Change in level = ", round(cf[3,1],r) , " (", round(CI[3,1],r), ", ", round(CI[3,2],r), "), p=", round(cf[3,4],r), "\n",
               "Change in slope = ",  round(cf[4,1],r) , " (", round(CI[4,1],r), ", ", round(CI[4,2],r), "), p=", round(cf[4,4],r) , sep="")
  ts.plot(plotTimeSeries, xlab=NULL,  ylab=ylab, ylim=ylim, main=main, sub=sub)
  title(xlab=xlab, line=5)

  usr <- par("usr")
  lines(rep(lastPreMonth[1] + lastPreMonth[2]/(frequency) ,10), seq(usr[3],usr[4],length.out=10), col="blue", main=NULL)
  ###  Replot data over hashing
  lines(plotTimeSeries , lty="solid", col="black", main=NULL)

  ### Add moving average
  movingAverage <- ts(f_mav(plotTimeSeries),start=firstMonth,
                      end=lastMonth,
                      frequency=frequency)
  lines(movingAverage, lty=1,col='purple', main=NULL)

  # fitted <- ts( its.lm$fitted, start=firstMonth,
  #               end=lastMonth,
  #               frequency=frequency)
  # fitted <- ts( c(its.lm$fitted[1:24], rep(NA,12), its.lm$fitted[25: length(its.lm$fitted)]), start=firstMonth,
  #               end=lastMonth,
  #               frequency=frequency)

  preT <- (2012-firstMonth[1])*12 + 8-firstMonth[2] + 1

  fitted <- ts( c(its.lm$fitted[1:preT], rep(NA,12), its.lm$fitted[(preT+1): length(its.lm$fitted)]), start=firstMonth,
                end=lastMonth,
                frequency=frequency)

  ## Add pre intervetion trend
  preVal <-   window(fitted, start=firstMonth,end=lastPreMonth)
  lines(preVal, lty= "dotted", col='green', lwd=4, main=NULL)

  ## Add post intervention trend
  postVal <- window(fitted, start=firstPostMonth,end=lastMonth)
  lines(postVal,  lty= "dotted", col='red', lwd=4, main=NULL)

  legend("topright", legend = c("Risk-adjusted", "Moving Average", "Fitted Pre", "Fitted Post"), col=c("black", "purple", "green", "red"), lty=c(1,1,3,3), cex=0.9)

  return(its.lm)
} # splitPlot <- function (



f_splitPlot2_glm <- function ( plotTimeSeries, switchObservation, frequency, firstMonth , lastPreMonth ,
                           firstPostMonth ,  lastMonth ,
                           xlab="Date", ylab="Count", main=NULL, sub=NULL, arOrder,
                           ylim=NULL, r) {
  ## Plot outline with shading
  require(forecast)
  library(nlme)
  base <-  switchObservation # row number for the first post period

  ## Intrupted time series model
  time <- 1:NROW(plotTimeSeries)
  intervention <- c(rep(0,base-1),rep(1,NROW(plotTimeSeries)-base+1))
  timeAfterIntervention  <- c(rep(0,base-1),1:((NROW(plotTimeSeries)-base+1)) )

  lmcoef <- coef(lm(plotTimeSeries ~ time + intervention +  timeAfterIntervention, data=patUse2, na.action="na.omit"))
  m <- length(lmcoef)

  its.lm <- glm(plotTimeSeries ~ time + intervention +  timeAfterIntervention,
              family=Gamma(link="log"),
              na.action="na.omit",
              start=c(20, rep(0,m-1)))

  cf <- summary(its.lm)
  print(cf)

  coef <- cf$coefficients[,1]
  coef <- 100*(exp(coef)-1)
  pv <- cf$coefficients[,4]
  CI <- confint(its.lm)
  print(CI)
  CIconv <- 100*(exp(CI)-1)
  print(CIconv)

  sub <- paste("Overall trend = ", round(coef[2],3), "% (", round(CIconv[2,1],3), ", ", round(CIconv[2,2],3),"), p=", round(pv[2],3), "\n",
               "Change in level = ", round(coef[3],3) , "% (", round(CIconv[3,1],3), ", ", round(CIconv[3,2],3),"), p=", round(pv[3],3), "\n",
               "Change in slope = ",  round(coef[4],3) , "% (", round(CIconv[4,1],3), ", ", round(CIconv[4,2],3),"), p=", round(pv[4],3), sep="")

  ts.plot(plotTimeSeries, xlab=NULL,  ylab=ylab, ylim=ylim, main=main, sub=sub)
  title(xlab=xlab, line=5)

  usr <- par("usr")
  lines(rep(lastPreMonth[1] + lastPreMonth[2]/(frequency) ,10), seq(usr[3],usr[4],length.out=10), col="blue")
  ###  Replot data over hashing
  lines(plotTimeSeries , lty="solid", col="black")

  ### Add moving average
  movingAverage <- ts(f_mav(plotTimeSeries),start=firstMonth,
                      end=lastMonth,
                      frequency=frequency)
  lines(  movingAverage, lty=1,col='purple')

  fitted <- ts( c(its.lm$fitted[1:24], rep(NA,12), its.lm$fitted[25: length(its.lm$fitted)]), start=firstMonth,
                end=lastMonth,
                frequency=frequency)

  ## Add pre intervetion trend
  preVal <-   window(fitted, start=firstMonth,end=lastPreMonth)
  lines(  preVal, lty= "dotted", col='green', lwd=4)

  ## Add post intervention trend
  postVal <- window(fitted, start=firstPostMonth,end=lastMonth)
  lines(  postVal,  lty= "dotted", col='red', lwd=4)

  legend("topright", legend = c("Risk-adjusted", "Moving Average", "Fitted Pre", "Fitted Post"), col=c("black", "purple", "green", "red"), lty=c(1,1,3,3), cex=0.9)

  return(its.lm)
}
```

Run ITS model with marginal effects from the patient-level analysis
```{r}
# PACF order= 1 for all pats; =3 for surg, =5 for med pats

y <- c("tvdc_cpi")
y_ul <- c(4000)
y_lab <- c("Patient risk-adjusted total variable direct costs")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("~/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  print(col)
  rr <- f_splitPlot2(aggData.ts[,col],
                     switchObservation=32,
                     frequency = 12,
                     firstMonth = c(2011,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,12),
                     arOrder=1,
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     r=round
  )
  dev.off()
}

y <- c("tvdc_cpi_surg1")
y_ul <- c(50000)
y_lab <- c("Patient risk-adjusted total variable direct costs \nSurgical patients")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("~/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  print(col)
  rr <- f_splitPlot2(aggData.ts[,col],
                     switchObservation=32,
                     frequency = 12,
                     firstMonth = c(2011,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,12),
                     arOrder=3,
                     ylim=c(0,y_ul[init]),
                     ylab=paste0("Patient risk-adjusted total variable direct costs"," ($)"),
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     r=round
  )
  dev.off()
}

y <- c("tvdc_cpi_surg0")
y_ul <- c(4000)
y_lab <- c("Patient risk-adjusted total variable direct costs \nMedical patients")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("~/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  print(col)
  rr <- f_splitPlot2(aggData.ts[,col],
                     switchObservation=32,
                     frequency = 12,
                     firstMonth = c(2011,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,12),
                     arOrder=5,
                     ylim=c(0,y_ul[init]),
                     ylab=paste0("Patient risk-adjusted total variable direct costs"," ($)"),
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     r=round
  )
  dev.off()
}
```


Run ITS model and save plots for sub-category costs - skip
```{r}
# covars <- c("CMI","surgicalSURG","elixsum","female1","ageGroup2(10,20]", "ageGroup2(20,30]", "ageGroup2(30,40]", "ageGroup2(40,50]", "ageGroup2(50,110]", "raceGroupBlack", "raceGroupOther race", "raceGroupUnknown","insMedicare MC","insMedicaid FFS","insMedicaid MC", "insSelf-pay", "insCommercial/Other ins")

aggData.ts2 <- aggData.ts

# statistically significant partial ACF of order 1: MRI_cpi, inhalation_cpi, OR_cpi, POT_cpi, Blood_cpi,

# statistically significant partial ACF of order 2: tvdc_cpi, postop_cpi, Laboratory_cpi, Radiology_cpi, ICU_cpi, Pharmacy_cpi, supplies_cpi, routine_cpi

# statistically significant partial ACF of order 3: implants_cpi



# order 1
y <- c("MRI_cpi_cmi", "inhalation_cpi_cmi", "OR_cpi_cmi", "POT_cpi_cmi", "Blood_cpi_cmi")
y_ul <- c(30, 100, 1500,100, 300)
y_lab <- c("MRI", "Inhalation therapy", "OR", "Physical/Occupational therapy", "Blood costs")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)
#
#   aggData.ts2[25:36, col] <- NA
#   
  print(col)
  rr <- f_splitPlot2(aggData.ts2[,col],
                     # cbind(aggData.ts[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     arOrder = 1,
                     r = round[init],
                     main=paste0("Interrupted Time Series Model\nInflation and DRG weight-adjusted ", y_lab[init])
  )
  dev.off()
}

# order 2
y <- c("tvdc_cpi_cmi", "postop_cpi_cmi", "Laboratory_cpi_cmi", "Radiology_cpi_cmi", "ICU_cpi_cmi", "Pharmacy_cpi_cmi", "supplies_cpi_cmi", "routine_cpi_cmi", "blood_qty_cmi")
y_ul <- c(8000, 500, 500,300, 1500, 1500, 500, 2000, 1)
y_lab <- c("Total variable direct costs", "Post Op/Stepdown", "Laboratory", "Radiology", "ICU/CCU", "Pharmacy", "Medical/Surgical supplies", "Routine care", "Blood units")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)
#
#   aggData.ts2[25:36, col] <- NA
#   
  print(col)
  rr <- f_splitPlot2(aggData.ts2[,col],
                     # cbind(aggData.ts[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     arOrder = 2,
                     r = round[init],
                     main=paste0("Interrupted Time Series Model\nInflation and DRG weight-adjusted ", y_lab[init])
  )
  dev.off()
}

# order 3
y0 <- c("Implants")
y <- paste0(y0,"_cpi_cmi")
y_ul <- c(1000)
y_lab <- c("Implants")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  # aggData.ts2[25:36, col] <- NA

  rr <- f_splitPlot2(aggData.ts2[,col],
                     # cbind(aggData.ts2[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     main=paste0("Interrupted Time Series Model\nInflation and DRG weight-adjusted ", y_lab[init]),
                     arOrder=3,
                     r=round[init]
  )
  dev.off()
}



# statistically significant partial ACF of order 1: routine_cpi, MRI_cpi, inhalation_cpi, OR_cpi, POT_cpi, supplies_cpi, ICU_cpi, Blood_cpi, Pharmacy_cpi
y0 <- c("Pharmacy", "MRI", "Blood", "ICU", "OR", "supplies", "inhalation", "POT","routine")
y <- paste0(y0,"_cpi")
summary(aggData[y])
y_ul <- c(1500,  30, 300, 1500, 1500, 1000, 300, 100, 2000)
y_lab <- c("Pharmacy", "MRI", "Blood", "ICU/CCU", "OR", "Medical/Surgical supplies", "Inhalation therapy","Physical/Occupational therapy","Routine care")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  # aggData.ts2[25:36, col] <- NA

  print(col)
  rr <- f_splitPlot2(aggData.ts2[,col],
                     cbind(aggData.ts2[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     arOrder=1,
                     r= round[init]
  )
  dev.off()
}


# statistically significant partial ACF of order 2: postop_cpi, Laboratory_cpi, Radiology_cpi
y0 <- c("postop", "Laboratory", "Radiology")
y <- paste0(y0,"_cpi")
summary(aggData[y])
y_ul <- c(500, 500, 300)
y_lab <- c("Post Op/Stepdown","Laboratory", "Radiology")
round = rep(3, length(y))

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  # aggData.ts2[25:36, col] <- NA

  rr <- f_splitPlot2(aggData.ts2[,col],
                     cbind(aggData.ts2[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0,y_ul[init]),
                     ylab=paste0(y_lab[init]," ($)"),
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     arOrder=2,
                     r=round[init]
  )
  dev.off()
}

# statistically significant partial ACF of order 3: implants_cpi

```


LOS & Mortality outcomes

```{r}
y <- c("los", "death")
for (col in y) {
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/acf_",col,".pdf", sep=""))
  par(mfrow=c(2,1))
  acf(aggData.ts[,col], na.action = na.pass )
  pacf(aggData.ts[,col], na.action = na.pass )
  dev.off()
}
# order 2 : LOS, death

# confint(its)

y_lab <- c("Length of stay (Days)", "In-hospital mortality rate")

y_ul <- c(10,0.05)
round <- c(3,4)

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)
  print(y_ul[init])
  rr <- f_splitPlot2(aggData.ts2[,col],
                     # cbind(aggData.ts2[,covars]),
                     switchObservation=45,
                     frequency = 12,
                     firstMonth = c(2010,9),
                     lastPreMonth = c(2014,4),
                     firstPostMonth = c(2014,5),
                     lastMonth=c(2017,5),
                     ylim=c(0, y_ul[init]),
                     ylab=y_lab[init],
                     main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                     arOrder=2,
                     r=round[init]
  )
  dev.off()
}
```

Readmission outcome
```{r}
outcome <- ts(aggData.ts[17:81,c("readmit","death")], start=c(2012,1), end=c(2017,5), frequency = 12)

y <- c("readmit")
y_lab <- c("Hospital readmission rate")

for (col in y) {
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/acf_",col,".pdf", sep=""))
  par(mfrow=c(2,1))
  acf(outcome[, col], na.action = na.pass )
  pacf(outcome[, col], na.action = na.pass )
  dev.off()
}

y_ul <- c(0.5)
round <- c(4)

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  rr <- f_splitPlot2(outcome[,col],
                      # cbind(outcome[,covars]),
                      switchObservation=29,
                      frequency = 12,
                      firstMonth = c(2012,1),
                      lastPreMonth = c(2014,4),
                      firstPostMonth = c(2014,5),
                      lastMonth=c(2017,5),
                      ylim=c(0,y_ul[init]),
                      ylab=y_lab[init],
                      main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                      arOrder=1,
                      r=round[init]
  )
  dev.off()
}
```

O/E outcome
```{r}
patUse3 <- patUse2[patUse2$DischargeDateMonth >= "201309" & patUse2$explos > 0,]

patUse3$tvdc_cpi_cmi <- patUse3$tvdc_cpi / patUse3$CMI

test <- model.matrix(~0 + season + CMI + surgical + female + AgeGroup + ageGroup2 + raceGroup + ins, data=patUse3)

# mean of CPI-adjusted costs?
aggData <- ddply(cbind(patUse3, test), .(DischargeDateMonth), numcolwise(mean, na.rm=TRUE))

# create the monthly time-series data with mean outcome variables and covariates
aggData.ts <- ts(aggData, start=c(2013,9), end=c(2017,5), frequency = 12)

y <- c("oelos", "oelos_gt15")
y_lab <- c("Observed/Expected (O/E) LOS ratio", "Probability of O/E LOS ratio > 1.5")

for (col in y) {
  pdf(paste("~/Dropbox/Research/VBM/results/acf_",col,".pdf", sep=""))
  par(mfrow=c(2,1))
  acf(aggData.ts[,col], na.action = na.pass )
  pacf(aggData.ts[,col], na.action = na.pass )
  dev.off()
}

# O/E ratio has order 1 PACF
# O/E ratio > 1.5 has no autocorr


f_splitPlot2 <- function ( plotTimeSeries, switchObservation, frequency, firstMonth , lastPreMonth ,
                           firstPostMonth ,  lastMonth ,
                           xlab="Date", ylab="Count", main=NULL, sub=NULL, arOrder,
                           ylim=NULL, r) {
  ## Plot outline with shading
  require(forecast)
  library(nlme)
  base <-  switchObservation # row number for the first post period

  ## Intrupted time series model
  time <- 1:NROW(plotTimeSeries)
  intervention <- c(rep(0,base-1),rep(1,NROW(plotTimeSeries)-base+1))
  timeAfterIntervention  <- c(rep(0,base-1),1:((NROW(plotTimeSeries)-base+1)) )
  its.lm <- gls(plotTimeSeries ~ time + intervention +  timeAfterIntervention ,
                correlation=corARMA(p=arOrder), method="ML",
                na.action="na.omit")    
  cf1 <- summary(its.lm)
  cf <- cf1$tTable
  print(cf)

  CI <- confint(its.lm)
  print(CI)

  sub <- paste("Overall trend = ", round(cf[2,1],r) , " (", round(CI[2,1],r), ", ", round(CI[2,2],r), "), p=", round(cf[2,4],r), "\n",
               "Change in level = ", round(cf[3,1],r) , " (", round(CI[3,1],r), ", ", round(CI[3,2],r), "), p=", round(cf[3,4],r), "\n",
               "Change in slope = ",  round(cf[4,1],r) , " (", round(CI[4,1],r), ", ", round(CI[4,2],r), "), p=", round(cf[4,4],r) , sep="")
  ts.plot(plotTimeSeries, xlab=NULL,  ylab=ylab, ylim=ylim, main=main, sub=sub)
  title(xlab=xlab, line=5)

  usr <- par("usr")
  lines(rep(lastPreMonth[1] + lastPreMonth[2]/(frequency) ,10), seq(usr[3],usr[4],length.out=10), col="blue", main=NULL)
  ###  Replot data over hashing
  lines(plotTimeSeries , lty="solid", col="black", main=NULL)

  ### Add moving average
  movingAverage <- ts(f_mav(plotTimeSeries),start=firstMonth,
                      end=lastMonth,
                      frequency=frequency)
  lines(  movingAverage, lty=1,col='purple', main=NULL)

  fitted <- ts( its.lm$fitted, start=firstMonth,
                end=lastMonth,
                frequency=frequency)

  ## Add pre intervetion trend
  preVal <-   window(fitted, start=firstMonth,end=lastPreMonth)
  lines(  preVal, lty= "dotted", col='green', lwd=4, main=NULL)

  ## Add post intervention trend
  postVal <- window(fitted, start=firstPostMonth,end=lastMonth)
  lines(  postVal,  lty= "dotted", col='red', lwd=4, main=NULL)

  legend("topright", legend = c("CPI-adjusted", "Moving Average", "Fitted Pre", "Fitted Post"), col=c("black", "purple", "green", "red"), lty=c(1,1,3,3), cex=0.9)

  return(its.lm)
} # splitPlot <- function (



f_splitPlot2_lm <- function ( plotTimeSeries, switchObservation, frequency, firstMonth , lastPreMonth ,
                           firstPostMonth ,  lastMonth ,
                           xlab="Date", ylab="Count", main=NULL, sub=NULL, arOrder,
                           ylim=NULL, r) {
  ## Plot outline with shading
  require(forecast)
  library(nlme)
  base <-  switchObservation # row number for the first post period

  ## Intrupted time series model
  time <- 1:NROW(plotTimeSeries)
  intervention <- c(rep(0,base-1),rep(1,NROW(plotTimeSeries)-base+1))
  timeAfterIntervention  <- c(rep(0,base-1),1:((NROW(plotTimeSeries)-base+1)) )

  its.lm <- lm(plotTimeSeries ~ time + intervention +  timeAfterIntervention ,
              na.action="na.omit")

  cf <- summary(its.lm)
  print(cf)

  coef <- 100*cf$coefficients[,1]
  # coef <- 100*(exp(coef)-1)
  pv <- cf$coefficients[,4]
  CI <- confint(its.lm)
  print(CI)
  CIconv <- 100*CI
  print(CIconv)

  sub <- paste("Overall trend = ", round(coef[2],3), "pp (", round(CIconv[2,1],3), ", ", round(CIconv[2,2],3),"), p=", round(pv[2],3), "\n",
               "Change in level = ", round(coef[3],3) , "pp (", round(CIconv[3,1],3), ", ", round(CIconv[3,2],3),"), p=", round(pv[3],3), "\n",
               "Change in slope = ",  round(coef[4],3) , "pp (", round(CIconv[4,1],3), ", ", round(CIconv[4,2],3),"), p=", round(pv[4],3), sep="")

  ts.plot(plotTimeSeries, xlab=NULL,  ylab=ylab, ylim=ylim, main=main, sub=sub)
  title(xlab=xlab, line=5)

  usr <- par("usr")
  lines(rep(lastPreMonth[1] + lastPreMonth[2]/(frequency) ,10), seq(usr[3],usr[4],length.out=10), col="blue")
  ###  Replot data over hashing
  lines(plotTimeSeries , lty="solid", col="black")

  ### Add moving average
  movingAverage <- ts(f_mav(plotTimeSeries),start=firstMonth,
                      end=lastMonth,
                      frequency=frequency)
  lines(  movingAverage, lty=1,col='purple')

  fitted <- ts( its.lm$fitted, start=firstMonth,
                end=lastMonth,
                frequency=frequency)

  ## Add pre intervetion trend
  preVal <-   window(fitted, start=firstMonth,end=lastPreMonth)
  lines(  preVal, lty= "dotted", col='green', lwd=4)

  ## Add post intervention trend
  postVal <- window(fitted, start=firstPostMonth,end=lastMonth)
  lines(  postVal,  lty= "dotted", col='red', lwd=4)

  legend("topright", legend = c("CPI-adjusted", "Moving Average", "Fitted Pre", "Fitted Post"), col=c("black", "purple", "green", "red"), lty=c(1,1,3,3), cex=0.9)

  return(its.lm)
}


y <- c("oelos")
y_lab <- c("Observed/Expected (O/E) LOS ratio")
y_ul <- c(1.5)
round = c(3)

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  rr <- f_splitPlot2(aggData.ts[,col],
                      # cbind(aggData.ts[,covars]),
                      switchObservation=9,
                      frequency = 12,
                      firstMonth = c(2013,9),
                      lastPreMonth = c(2014,4),
                      firstPostMonth = c(2014,5),
                      lastMonth=c(2017,5),
                      ylim=c(0,y_ul[init]),
                      ylab=y_lab[init],
                      main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                      arOrder=1,
                      r=round[init]
  )
  dev.off()
}



y <- c("oelos_gt15")
y_lab <- c("Proportion of discharges having O/E LOS ratio > 1.5")
y_ul <- c(1)
round = c(3)

init <- 0
for (col in y) {
  init <-  init + 1
  pdf(paste("/Users/kunheekim/Dropbox/Research/VBM/results/its_",col,".pdf", sep=""))
  rr <- paste0("its_",col)

  rr <- f_splitPlot2_lm(aggData.ts[,col],
                      # cbind(aggData.ts[,covars]),
                      switchObservation=9,
                      frequency = 12,
                      firstMonth = c(2013,9),
                      lastPreMonth = c(2014,4),
                      firstPostMonth = c(2014,5),
                      lastMonth=c(2017,5),
                      ylim=c(0,y_ul[init]),
                      ylab=y_lab[init],
                      main=paste0("Interrupted Time Series Model", "\n", y_lab[init]),
                      r=round[init]
  )
  dev.off()
}
```

### For the explanation of the ITS model, make graphs - skip
```{r}
patUse3 <- subset(patUse2, DischargeDateD < "2012-09-01" | DischargeDateD > "2013-08-31" )

# CMI.lm <- lm(CMI ~ t + tpostInt + season, data=patUse3)
tv.glm <- glm(tvdc_cpi / CMI ~ t + tpostInt + int ,
              data = patUse3,
              na.action="na.omit")
# tv.glm <- glm(tvdc_cpi ~ t + tpostInt + int + CMI
#                 + surgical + ynel1 + ynel2 + ynel3 + ynel4 + ynel5 + ynel6 + ynel7 + ynel8 + ynel9 + ynel10 + ynel11 + ynel12 + ynel13 + ynel14 + ynel15 + ynel16 + ynel17 + ynel18 + ynel19 + ynel20 + ynel21 + ynel22 + ynel23 + ynel24 + ynel25 + ynel26 + ynel27 + ynel28 + ynel29 + ynel30 + ynel31 + female + AgeGroup + raceGroup + ins,
#               data = patUse3,
#               na.action="na.omit")
summary(tv.glm )
# confint(CMItv.ls )

fit.ls <- data.frame(y= tv.glm$y,
                     t=tv.glm$data$t,
                     yhat=tv.glm$fitted.values)
df <- ddply(fit.ls, .(t), summarise, yM = mean(y, na.rm=TRUE),  yMed = median(y, na.rm=TRUE), yhatM = mean(yhat, na.rm=TRUE),  yhatMed = median(yhat, na.rm=TRUE))
df <- melt(df, id.var="t")
df$variable <- factor(df$variable, labels = c("Mean total cost", "Median total cost", "Mean predicted total cost", "Median predicted total cost"))

# only raw mean
ggplot(df[df$variable=="Mean total cost",], aes(x=t, y=value, colour=variable, group=variable)) +
  geom_line() +
  labs(x="Month index", y="Mean total variable direct costs ($) per discharge",
       title="Mean inflation and CMI-adjusted actual total variable direct costs ($) \nper discharge by month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=45), colour="red") +
  ylim(0,7000)
ggsave("~/Dropbox/Research/VBM/results/ex_tvdc_cpi.pdf", width=20, height=10, units="cm")

# only raw & predicted mean
ggplot(df[df$variable=="Mean total cost" | df$variable=="Mean predicted total cost",], aes(x=t, y=value, colour=variable, group=variable)) +
  geom_line() +
  labs(x="Month index", y="Mean total variable direct costs ($) per discharge",
       title="Mean inflation and CMI-adjusted actual and predicted \ntotal variable direct costs ($) per discharge by month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=45), colour="red") +
  ylim(0,7000)
ggsave("~/Dropbox/Research/VBM/results/ex_tvdc_cpi2.pdf", width=20, height=10, units="cm")



sb <- summaryBy(y~ t,data=fit.ls,  FUN=c(mean, median))
sc <- summaryBy(yhat~ t,data=fit.ls,  FUN=c(mean, median))

plot(sb$t, sb$y.mean, col="black", type="l", ylim=c(0,7000))
lines(sc$t, sc$yhat.mean, col="black", lty=2)
abline(v=45)
```



### Plot fitted values vs observed values - SKIP?

```{r}
y <- c("tvdc_cpi")
comb <- c("t + int + tpostInt + season + cmiGroup + surgical + female + AgeGroup + raceGroup + ins")

for (col in y) {
  sp <- paste0(col,"~", comb)
  lmcoef <- coef(lm(sp, data=patUse2, na.action="na.omit"))
  k <- length(lmcoef)

  model <- glm(sp,
      data = patUse2,
      family=Gamma(link="log"),
      na.action="na.omit",
      start=c(10, rep(0,k-1)))
}

patUse2$postFitAll <- predict(model, newdata = patUse2,family = Gamma(link="log"), type = "response")
summary(postFitAll)

# postFitInt <- predict(model, newdata = patUse2[patUse2$int==1,],family = Gamma(link="log"), type = "link")

plotPatLevel <- data.frame(fit=postFitAll,
  # fit=model$fitted.values,
                           y=patUse2$tvdc_cpi,
                           t=model$data$t)

library(doBy)
fitData <- summaryBy(fit~t, data=plotPatLevel, FUN=c(median,mean))
yData <- summaryBy(y~t, data=plotPatLevel, FUN=c(median,mean))

# add in rows for missing FY


plot(yData$t, yData$y.median, col="red", type="l", ylim=c(0, 12000))
lines(fitData$t, fitData$fit.median, col="blue", lty=2)

plot(yData$t, yData$y.mean, col="red", type="l", ylim=c(0, 12000))
lines(fitData$t, fitData$fit.mean, col="blue", lty=2)

plot(yData$t, yData$y.median, col="red", type="l", ylim=c(0, 15000))
lines(fitData$t, fitData$fit.median, col="blue", lty=2)

ggplot(patUse2, aes(y=postFitAll, x=tvdc_cpi)) +
  geom_point()

```


# Diff-in-diff analysis using the quarterly COTH data

```{r}
# pre 2012 data
cothpre <- read_excel("~/Box Sync/VBM/Data/Copy of Historical COTH Data_2009-2011.xlsx")
cothpre <- cothpre[2:nrow(cothpre),]
cothpre <- t(cothpre)
colnames(cothpre) <- c("qtr","NYU",  "Median","P75", "P25" )
cothpre <- cothpre[c("qtr", "NYU", "P75", "Median", "P25")]
# cothpre <- cothpre[NROW(cothpre),]
a <- cothpre[13,]
a[1] <- "2011 Q4"

# post 2012 data
coth <- read_excel("~/Box Sync/VBM/Data/coth_wage_cmi_adj_expense_perdisch.xlsx")
# coth <- read.csv("~/Box Sync/VBM/Data/coth_wage_cmi_adj_expense_perdisch.csv")
colnames(coth) <- c("qtr", "NYU", "P75", "Median", "P25" )

# # add 2011 Q4 row
# coth <- as.data.frame(rbind(a, as.matrix(coth)))

coth$post <- c(rep(0,10),rep(1,nrow(coth)-10))

# suppress FY 2013
coth[4:7,2:6] <- NA

# library(zoo)
# coth$qtr <- as.yearqtr(coth$qtr, format = "%Y Q%q")

# reshape to long data
coth.l <- reshape(coth,
                  varying = c("NYU", "P75", "Median", "P25"),
                  v.names = "cost",
                  timevar = "gp",
                  times = c("NYU", "P75", "Median", "P25"),
                  direction = "long")

# create NYU hosp indicator
coth.l$nyu <- ifelse(coth.l$gp=="NYU",1,0)

#destring cost variable
coth.l$cost <- as.numeric(gsub(",", "", coth.l$cost))
```

plot raw data for different groups
```{r}
coth.l$gp <- factor(coth.l$gp, levels=c("NYU", "P25", "Median","P75"))
# coth.l$gp[is.na(coth.l$cost)] <- NA
ddply(coth.l, .(gp), summarise, M=mean(cost, na.rm=TRUE))

ggplot(coth.l) +
  geom_line(aes(x=qtr, y=cost, colour=gp, group=gp)) +
  theme_grey(base_size=13) +
  theme(axis.text.x = element_text(angle=45, vjust = 0.5),
        legend.title=element_blank(),
        legend.position="bottom") +
  scale_y_continuous(limits= c(0,12000), minor_breaks = seq(0,12000,1000)) +
  labs(title="Wage and CMI Standardized Expense per Adjusted Discharge \nNYU vs 25th, 50th, 75th percentiles of all teaching hospitals",
       y="Expense per discharge",
       x="Year-quarter")
ggsave("~/Dropbox/Research/VBM/results/raw_coth.pdf", width=20, height=10, units="cm")
```

Diff-in-Diff using post indicator
```{r}
did <- lm(cost ~ post + nyu + post*nyu,
          data=test)
summary(did)

pct_chng <- function(x) 100*(exp(x)-1)

for (i in c("P25", "Median", "P75")) {
  test <- coth.l[coth.l$gp=="NYU" | coth.l$gp==i, ]
  # did <- lm(cost ~ post + nyu + post*nyu,
  #         data=test)
  # print(summary(did))

  x <- paste0("did.glm_",i)
  print(x)
  x <- glm(formula = cost ~ nyu + post + post*nyu,
              data = test,
              family=Gamma(link="log"),
              na.action="na.omit")
  print(summary(x))

  stargazer(x, type="html", report = "vcp*",
          t.auto=F, p.auto=F,
          # keep.stat=c("n","adj.rsq","ll","aic"),
          apply.coef = pct_chng,
          dep.var.labels = depv,
          covariate.labels = c("Post","NYU","NYU X Post"),
          out=paste("~/Dropbox/Research/VBM/results/did.glm_", i,".htm", sep=""))
}
```



get quarterly coefficients
```{r}
# regression on NYU indicator
test <- coth.l[coth.l$gp=="NYU" | coth.l$gp=="P25" | coth.l$gp=="Median", ]
did <- lm(cost ~ qtr + nyu + qtr*nyu,
          data=coth.l)
summary(did)

did.glm <- glm(formula = cost ~ qtr + nyu + qtr*nyu,
              data = coth.l,
              family=Gamma(link="log"),
              na.action="na.omit")
summary(did.glm)
CI <- confint(did.glm)
CI2 <- 100*(exp(CI)-1)
coeff_pc <- 100*(exp(did.glm$coefficients)-1)

didcoef <- cbind(coeff_pc, CI2)
didcoef <- didcoef[24:44,]
didcoef <- as.data.frame(didcoef)

# create qtr var
didcoef$qtr_n <- c(2:4,rep(1:4,4),1:2)
didcoef$qtr_yr <- c(rep(2012,3), rep(2013,4), rep(2014,4), rep(2015,4), rep(2016,4), rep(2017,2))

library(zoo)
didcoef$qtr <- as.yearqtr(paste(didcoef$qtr_yr, didcoef$qtr_n, sep=" Q"), format = "%Y Q%q")

# suppress FY 2013
didcoef$coeff_pc[2:5] <- NA
didcoef$`2.5 %`[2:5] <- NA
didcoef$`97.5 %`[2:5] <- NA

# didcoef$id <- 1:NROW(didcoef)
# ggplot(didcoef) +
#   geom_point(aes(x=id, y=coeff_pc)) +
#   geom_errorbar(aes(x=id, ymin=`2.5 %`, ymax=`97.5 %`)) +
#   scale_y_continuous(limits = c(-50,100))

ggplot(didcoef) +
  geom_point(aes(x=qtr, y=coeff_pc)) +
  geom_errorbar(aes(x=qtr, ymin=`2.5 %`, ymax=`97.5 %`)) +
  geom_vline(aes(xintercept=2014.25), colour="red") +
  scale_y_continuous(limits = c(-50,100),
                     minor_breaks=seq(-50,100,10)) +
  scale_x_continuous(breaks=seq(2012.25, 2017.25, 0.25),
                     labels=paste(didcoef$qtr_yr, didcoef$qtr_n, sep=" Q")) +
  labs(x="Year-Quarter",
       y="Percentage change (%)",
       title="Wage and DRG weight standardized expense per adjusted discharge \nin NYU vs 25th, 50th, 75th percentiles for all other teaching hospitals") +
  theme_grey(base_size=13) +
  theme(axis.text.x = element_text(angle=45, vjust = 0.5))
ggsave("~/Dropbox/Research/VBM/results/did_coth.pdf", width=20, height=10, units="cm")
```



# Diff-in-diff analysis using hospital-quarter level Vizient data

## prepare Vizient data
import Vizient data
```{r}
allth_cmi1 <- read_excel("~/Box Sync/VBM/Data/Vizient (lucy)/NYU Historical Data Request v2.xlsx", sheet = "CMI", skip=34, col_names = T)
allth_cmi2 <- read_excel("~/Box Sync/VBM/Data/Vizient (lucy)/Historical 2014 CMI - Cost by Discharge Quarter.xlsx", sheet = "CMI", skip=36, col_names = T)
allth_cmi3 <- read_excel("~/Box Sync/VBM/Data/Vizient (lucy)/Viz_cmi_2014-4_2017-4.xlsx", skip=36, col_names = T)

allth_cost1 <- read_excel("~/Box Sync/VBM/Data/NYU Historical Data Request v2.xlsx", sheet = "Cost", skip=34, col_names = T)
myvars <- names(allth_cost1) %in% c("X__1") 
allth_cost1 <- allth_cost1[!myvars] 
allth_cost2 <- read_excel("~/Box Sync/VBM/Data/Vizient (lucy)/Historical 2014 CMI - Cost by Discharge Quarter.xlsx", sheet = "Cost", skip=36, col_names = T)
allth_cost3 <- read_excel("~/Box Sync/VBM/Data/Vizient (lucy)/Viz_cost_2014-4_2017-4.xlsx", skip=36, col_names = T)
```

append data
```{r}
cmidf <- rbind(allth_cmi1, allth_cmi2, allth_cmi3)
costdf <- rbind(allth_cost2, allth_cost3)
costdf <- costdf[order(costdf$Hospital, costdf$`Discharge Quarter`),]
# write_csv(costdf, "~/Box Sync/VBM/Data/Vizient (lucy)/Viz_costdf.csv")
# write_csv(costdf, "~/Box Sync/VBM/Data/Vizient (lucy)/Viz_cmidf.csv")

myvars <- names(costdf) %in% c("Direct Cost StatSig") 
costdf <- costdf[!myvars] 
costdf <- rbind(costdf, allth_cost1)

# in cost data, replace the period 2012-4 to 2013-3 with the newly received data applying the "2015 Model"
fy13 <- read.csv("~/Box Sync/VBM/Data/Vizient (lucy)/CDB Historical Cost Archive Runs 2012-4 to 2013-3.csv", skip=36)
fy13 <- fy13[c(2,3,5:10,12:15)]
names(fy13)[1] <- "Discharge Quarter"
names(fy13)[2] <- "Hospital"
names(fy13)[4] <- "LOS Outliers"
names(fy13)[5] <- "Direct Cost Cases Reported"
names(fy13)[6] <- "LOS Outliers Reported"
names(fy13)[7] <- "Mean Total Cost (Obs)"
names(fy13)[8] <- "Mean Direct Cost (Obs)"
names(fy13)[9] <- "StDev Direct Cost (Obs)"
names(fy13)[10] <- "Mean Direct Cost (Exp)"
names(fy13)[11] <- "Direct Cost Index"
names(fy13)[12] <- "Direct Cost Variance ($M)"
costdf2 <- subset(costdf, `Discharge Quarter`!="2012-4" & `Discharge Quarter`!="2013-1" & `Discharge Quarter`!="2013-2" & `Discharge Quarter`!="2013-3")
costdf3 <- rbind(costdf2, fy13)

# merge cost & CMI data
df <- merge(costdf3, cmidf, by=c("Hospital","Discharge Quarter"))
table(df$`Discharge Quarter`)

#restrict to FY 14-16: 2013/09 - 2017/05 
df <- subset(df, `Discharge Quarter`!="2011-4" & `Discharge Quarter`!="2012-1" & `Discharge Quarter`!="2012-2" & `Discharge Quarter`!="2012-3" & `Discharge Quarter`!="2012-4" & `Discharge Quarter`!="2013-1" & `Discharge Quarter`!="2013-2" & `Discharge Quarter`!="2013-3")
table(df$`Discharge Quarter`)

# subset data to relevant variables
v <- c("Hospital", "Discharge Quarter","Mean Total Cost (Obs)","Mean Direct Cost (Exp)","CMI", "Cases.x", "Cases.y", "Direct Cost Cases Reported")
df <- df[v]
colnames(df)[3] <- "mtcost"
colnames(df)[4] <- "mdcost"
colnames(df)[2] <- "qtr"
colnames(df)[6] <- "cost_cases"
colnames(df)[7] <- "CMI_cases"
colnames(df)[8] <- "dcost_cases"

vizdf <- df

# create NYU hosp indicator
vizdf$nyu <- ifelse(vizdf$Hospital=="330214 NYU",1,0)

# extract Medicare provider # 
vizdf$provid <- as.numeric(substr(vizdf$Hospital, start = 1, stop = 6))

library(zoo)
vizdf$qt <- as.yearqtr(vizdf$qtr, format = "%Y-%q")
table(vizdf$qt)
```

drop if hospitals have 0 cost in some quarters - maybe closed; no hospital had 0 cost - SKIP
```{r}
# vizdf <- data.table(vizdf)
# vizdf2 <- vizdf[, .SD[which.min(mdcost)!=0], by=Hospital]
# 
# test <- vizdf[, .SD[which.min(mdcost)], by=provid]
# test <- as.data.frame(test[mdcost==0,provid])
# test$x <- 1:NROW(test)
# colnames(test)[1] <- "provid"
# 
# vizdf2 <- merge(test,vizdf, by=c("provid"), all.y=T)
# vizdf2 <- subset(vizdf2, is.na(x)==T)
# vizdf <- vizdf2
# 
# table(vizdf$qt)
```

create a balanced panel
```{r}
# how many hospitals have a balaned panel?
length(table(vizdf$qt))
test <- ddply(vizdf, .(provid), summarise, Nqtr=length(provid))
table(test$Nqtr) # 65 hospitals stayed throughout out of 95
sum(table(test$Nqtr))

# restrict to balanced panel 
vizdf <- merge(test,vizdf, by=c("provid"), all.y=T)
vizdf <- subset(vizdf, Nqtr==length(table(vizdf$qt)))
rm(test)

table(vizdf$qt)
# 139 hospitals/ qtr

vizdf$mdcost <- as.numeric(vizdf$mdcost)
vizdf$mtcost <- as.numeric(vizdf$mtcost)
vizdf$dcost_cases <- as.numeric(vizdf$dcost_cases)
```

check the cost & # cases trend for non-NYU hospitals
```{r}

# all other teaching hospitals
df4 <- subset(vizdf, nyu==0)
keep <- c("mdcost", "provid", "qtr", "dcost_cases")
test <- ddply(df4[keep], .(qtr), numcolwise(mean, na.rm=TRUE))

# line graph of CMI & inflation adjusted cost at the quarter level just for NYU 
ggplot(test, aes(x=qtr, y=mdcost, group=1)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Mean direct cost ($)",
       title="Raw mean direct cost ($) by quarter",
       subtitle="All Other Teaching Hospitals",
       caption="Data source: 139 hospitals that had non-zero mean direct cost throughout 2013 Q4 - 2017 Q4"
       ) +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,11000),
                     breaks=seq(0,11000,2000)) +
  scale_colour_discrete(guide = guide_legend(ncol=1))
ggsave("~/Dropbox/Research/VBM/results/mdcost_trajectory_raw_other.pdf", width=20, height=10, units="cm")

# plot # cases
ggplot(test, aes(x=qtr, y=dcost_cases, group=1)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Number of cases",
       title="Number of cost cases by quarter",
       subtitle="All Other Teaching Hospitals",
       caption="Data source: 139 hospitals that had non-zero mean direct cost throughout 2013 Q4 - 2017 Q4") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,10000),
                     breaks=seq(0,10000,2000)) 
ggsave("~/Dropbox/Research/VBM/results/cost_cases_trajectory_other2.pdf", width=20, height=10, units="cm")

```

## prepare NYU data

construct the mean direct cost & mean total cost per quarter for NYU using the internal data applying the same restrictions as when extracting Vizient data
(copy and pasted from above)
```{r}
load("~/Box Sync/VBM/Data/vbmR/sample-full2017-postMay2014.RData")

#subset the first data to Sep 2011 - Aug 2016 (exclude FY 11 & 2017)
pat3 <- subset(pat, pat$DischargeDateD >= "2011-09-01" & pat$DischargeDateD < "2016-09-01")
min(pat3$DischargeDateD ) # "2011-09-01"
max(pat3$DischargeDateD ) #  "2016-08-31"

# before appending data, remove problematic variables
v <- names(pat3) %in% c("NYU Reporting Department – Most Recent", "NYU Reporting Division – Most Recent")
pat3 <- pat3[!v]
v <- names(pat2) %in% c("NYU Reporting Department – Most Recent", "NYU Reporting Division – Most Recent")
pat2 <- pat2[!v]
pat3$`Admit Service` <- as.numeric(pat3$`Admit Service`)
pat3$`Admit Source` <- as.numeric(pat3$`Admit Source`)
pat3$`Admission Type` <- as.numeric(pat3$`Admission Type`)
pat3$`EPIC Discharge Dept` <- as.numeric(pat3$`EPIC Discharge Dept`)
pat3$`Discharge Service` <- as.numeric(pat3$`Discharge Service`)
pat3$`Discharge Disposition` <- as.numeric(pat3$`Discharge Disposition` )
pat3$`MS-DRG` <- as.numeric(pat3$`MS-DRG`)
pat3$`NY-DRG` <- as.numeric(pat3$`NY-DRG`)

# append the two datasets
patUse <- bind_rows(pat3, pat2, .id="id" )
table(patUse$id)
min(patUse$DischargeDateD ) # "2011-09-01"
max(patUse$DischargeDateD ) #  "2017-12-31"

patUse$DischargeDateMonth <- format(patUse$DischargeDateD  , "%Y%m")  
NROW(patUse)

# Exclude non-inpatient people
sum(is.na(patUse$PatientClass_Desc))
table(patUse$PatientClass_Desc)
patUse <- subset(patUse, PatientClass_Desc=="Inpatient")
NROW(patUse)

# exclude DRG weight = 0 or MS-DRG = 998 or 999
table(patUse$`MS DRG Weight (CMI)`)
patUse <- subset(patUse, `MS-DRG`!=999 & `MS-DRG`!=998 & `MS DRG Weight (CMI)`!=0)
summary(patUse$`MS DRG Weight (CMI)`)
table(patUse$`MS DRG Weight (CMI)`)
NROW(patUse)

# Exclude Rehab hospital & rehab discharges by using reporting department = Rehab
patUse <- subset(patUse, `NYU Reporting Department – Most Recent_Desc`!="Rehabilitation Medicine" & `MS-DRG Type`!="REHAB")
NROW(patUse)

# exclude MS-DRG type is missing and reporting department is missing
patUse <- patUse[complete.cases(patUse[, c("NYU Reporting Department – Most Recent_Desc", "MS-DRG Type")]),]

# exclude if reporting department is unknown
patUse <- subset(patUse, `NYU Reporting Department – Most Recent_Desc`!="UNK")

# exclude Lutheran
patUse <- subset(patUse, str_sub(patUse$`EPIC Discharge Dept_Desc`,1,1) != "L" | is.na(patUse$`EPIC Discharge Dept_Desc`))
NROW(patUse)

table(patUse$DischargeDateMonth)

# focus on inpatient cases in Tisch 
table(patUse$`Custom Patient Type`)
table(patUse$DischargeDateMonth, patUse$`Custom Patient Type`)
# HI patients : ~500 patients per month
patUse <- subset(patUse, !is.na(`Custom Patient Type`) & `Custom Patient Type` %in% c("TI"))

# exclude los outlier = mean + 3*SD (use 3 to be conservative)
mean(patUse$`Length of Stay`) + 3*sd(patUse$`Length of Stay`) #25 days
sum(patUse$`Length of Stay` > 25)
sum(patUse$`Length of Stay` > 25) / NROW(patUse)

patUse <- subset(patUse, `Length of Stay` <= 25 & `Length of Stay` >= 1)

# can't identify nonviable neonate
# can't identify hospice patients

# create quarter 
patUse$yr <- year(patUse$DischargeDateD)
patUse$mo <- month(patUse$DischargeDateD)
patUse$qt <- 1 + floor((patUse$mo-1) / 3 )
patUse$qtr <- paste0(patUse$yr, "-",patUse$qt)
library(zoo)
patUse$qt <- as.yearqtr(patUse$qtr, format = "%Y-%q")
table(patUse$qt)

# get 2 cost variables : mean direct costs & mean total costs per quarter: `Enc - Total Direct Costs`, `Enc - Total Indirect Costs`

# get quarterly mean costs & # cases
nyudf <- ddply(patUse, .(qt, qtr), summarize, N=length(`Enc - Total Direct Costs`), mdcost_nyu = mean(`Enc - Total Direct Costs`, na.rm=TRUE), mtcost_nyu = mean(`Enc - Total Direct Costs` + `Enc - Total Indirect Costs`, na.rm=TRUE), CMI_nyu = mean(`MS DRG Weight (CMI)`, na.rm=TRUE) )
```

merge the Vizient aggregate data with aggregate data constructed from NYU data
```{r}
df <- merge(vizdf, nyudf, by=c("qtr"))

library(zoo)
df$qt <- as.yearqtr(df$qtr, format = "%Y-%q")
table(df$qt)

# drop 2017 Q4 b/c of the drop in # cases per hospital for non-NYU hospitals (probably the methodology changes starting from 2017 Q4)
df <- subset(df, qtr!="2017-4")

df$cost_cases <- as.numeric(df$cost_cases)

```

are NYU costs comparable b/w Vizient and internal data
```{r}
test <- subset(df, provid==330214)

test$mdcost_diff <- (test$mdcost - test$mdcost_nyu)/test$mdcost
test$mtcost_diff <-( test$mtcost - test$mtcost_nyu)/test$mtcost
summary(test$mdcost_diff)
summary(test$mtcost_diff)
test$n_diff <- (test$cost_cases - test$N)/test$cost_cases
summary(test$n_diff)
diff <- (df$CMI - df$CMI_nyu)/df$CMI
summary(diff)
# Vizient costs are always 40-80% lower than aggregate from internal data -> internal costs are higher 
# Vizient CMI is higher -> internal CMI is lower


ggplot(test) +
  geom_line(aes(x=qtr, y=cost_cases, colour="Vizient", group=1)) +
  geom_line(aes(x=qtr, y=N, colour="Internal", group=1)) +
  geom_point(aes(x=qtr, y=cost_cases, colour="Vizient")) +
  geom_point(aes(x=qtr, y=N, colour="Internal")) +
  labs(x="quarter", 
       y="Number of cases",
       title="Number of cases for NYU hospital by data source") +
  theme(legend.title=element_blank(), legend.position = "bottom") +
  scale_y_continuous(limits = c(0,12000)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=11), colour="red") 
ggsave("~/Dropbox/Research/VBM/results/ncases_bydatasource.png")  


ggplot(test) +
  geom_line(aes(x=qt, y=CMI, colour="Vizient")) +
  geom_line(aes(x=qt, y=CMI_nyu, colour="Internal")) +
  geom_point(aes(x=qt, y=CMI, colour="Vizient")) +
  geom_point(aes(x=qt, y=CMI_nyu, colour="Internal")) +
  labs(x="quarter", 
       y="Mean DRG weights",
       title="Mean DRG weights for NYU hospital by data source") +
  theme(legend.title=element_blank(), legend.position = "bottom") +
  scale_y_continuous(limits = c(0,2))
ggsave("~/Dropbox/Research/VBM/results/nyuCMI_bydatasource2.png")  

ggplot(test) +
  geom_line(aes(x=qt, y=mdcost, colour="Vizient")) +
  geom_line(aes(x=qt, y=mdcost_nyu, colour="Internal")) +
  geom_point(aes(x=qt, y=mdcost, colour="Vizient")) +
  geom_point(aes(x=qt, y=mdcost_nyu, colour="Internal")) +
  labs(x="quarter", 
       y="Mean direct cost",
       title="Mean direct cost for NYU hospital by data source") +
  theme(legend.title=element_blank(), legend.position = "bottom") +
  scale_y_continuous(limits = c(0,20000))
ggsave("~/Dropbox/Research/VBM/results/nyumdcost_bydatasource.png")  

ggplot(test) +
  geom_line(aes(x=qt, y=mtcost, colour="Vizient")) +
  geom_line(aes(x=qt, y=mtcost_nyu, colour="Internal")) +
  geom_point(aes(x=qt, y=mtcost, colour="Vizient")) +
  geom_point(aes(x=qt, y=mtcost_nyu, colour="Internal")) +
  labs(x="quarter", 
       y="Mean total cost",
       title="Mean total cost for NYU hospital by data source") +
  theme(legend.title=element_blank(), legend.position = "bottom") +
  scale_y_continuous(limits = c(0,30000))
ggsave("~/Dropbox/Research/VBM/results/nyumtcost_bydatasource.png")  

```

import inflation data 
```{r}
# adjust for inflation 
infl <- read_excel("~/Box Sync/VBM/Data/medical_inflation_03102018.xlsx", range = cell_rows(12:132), col_names=TRUE) # seasonally adjusted

# create month 
infl$day <- rep(1,nrow(infl))
infl$Period <-gsub('M','', infl$Period)
infl$Period <- as.numeric(infl$Period)
infl$date <- paste(infl$Year, infl$Period, infl$day, sep="/")
infl$date <- as.Date(infl$date)

# get quarterly mean inflation rate
infl$yq <- as.yearqtr(infl$date, format = "%Y-%m-%d")

colnames(infl)[4] <- "cpi"

infl <- subset(infl, yq >= "2013 Q4" & yq <= "2017 Q3")

infl.q <- ddply(infl, .(yq), summarize, mcpi = mean(cpi))
```

merge inflation data with discharge-level data & adjust for inflation 
```{r}
df <- merge(df, infl.q, by.x="qt", by.y="yq")

# rebase using 2011 Q4 CPI
y <- c("mtcost", "mdcost", "mtcost_nyu", "mdcost_nyu")
for (col in y) {
  new.col <- paste(col, '_cpi', sep='')
  df[, new.col] <- df[, col] * 404.6390/df$mcpi
}

#create FY var := federal fiscal year ending in Sept
t <- as.data.frame(sort(unique(df$qt)))
t$i <- as.numeric(rownames(t))
t$fy <- 2014 + floor((t$i-1)/4)
colnames(t)[1] <- c("qt")
t <- t[c("qt","fy")]

df <- merge(df, t, by=c("qt"))

# # suppress raw data for FY 2013 only for NYU hospitals - irrelevant b/c the sample period starts from FY 14
# vars <- c("mdcost_cpi", "mdcost_nyu_cpi")
# df[df$qt >= "2012 Q4" & df$qt <= "2013 Q3" & df$provid==330214,vars] <- NA
# # df[df$qt >= "2012 Q4" & df$qt <= "2013 Q3" & df$provid==330214,11:13] <- NA

```


check that inflation adjustment was done well
```{r}
cc <- ddply(df, .(qt), summarise, mmdcost =mean(mdcost_nyu, na.rm=TRUE), mmdcost_cpi =mean(mdcost_nyu_cpi, na.rm=TRUE))

# create monthly time-series data
infl.ts <- ts(infl.q, start = c(2011,4), end=c(2017,2), frequency = 4)

cc.ts <- ts(cc, start = c(2011,4), end=c(2017,2), frequency = 4)
# monthly_cost.ts <- ts(monthly_cost[25:nrow(monthly_cost),], start = c(2013,9), end=c(2017,5), frequency = 12)

pdf("~/Dropbox/Research/VBM/results/inflation_check_viz.pdf")
par(mfrow=c(2,1))
ts.plot(infl.ts, ylab="CPI", ylim=c(0,500))
ts.plot(cc.ts[,2:3],
        gpars=list(xlab="Quarter", ylab="Mean Direct Cost ($)", lty=c(1:2)),
        ylim=c(0,20000))
legend("topright", legend = c("Unadjusted", "CPI-adjusted"), lty = 1:2, cex=.65)
dev.off()
```

get Medicare cost report data for hosp chars 
```{r}
cr <- read.csv("~/Box Sync/VBM/Data/hosp_cr_VBM.csv")
myvars <- names(cr) %in% c("provid", "fy", "teaching", "own_np", "own_fp", "own_gv", "uncomp1","dissh", "beds", "size", "urban") 
cr <- cr[myvars] 

# fill in missing data from the previous row by group
require(tidyverse) #fill is part of tidyr
cr <- cr[order(cr$provid, cr$fy),]
```

merge the Vizient data with CR data
```{r}
df2 <- merge(df, cr, by=c("provid","fy"))

# create the ITS variables: time, intervention, time after intervention
t <- as.data.frame(sort(unique(df2$qt)))
colnames(t)[1] <- c("qt")

# intervention occurred in Q2 2014
t$int <- ifelse(t$qt >= "2014 Q3",1,0) 
t$t <- 1:nrow(t)
lastpre <- 3
t$tpostInt <- c(rep(0,lastpre), 1:(nrow(t)-lastpre))

# merge with ITS variables
df2 <- merge(df2, t, by=c("qt"))

table(df$qt)
table(df2$qt)
```


```{r}
df3 <- df2

# create log cost
df3$lnmdcost_cpi <- log(df3$mdcost_cpi)
df3$lnmtcost_cpi <- log(df3$mtcost_cpi)

# for NYU, replace the costs & CMI with NYU internal data
df3$mdcost[df3$nyu==1] <- df3$mdcost_nyu[df3$nyu==1]
df3$mtcost[df3$nyu==1] <- df3$mtcost_nyu[df3$nyu==1]
df3$mdcost_cpi[df3$nyu==1] <- df3$mdcost_nyu_cpi[df3$nyu==1]
df3$mtcost_cpi[df3$nyu==1] <- df3$mtcost_nyu_cpi[df3$nyu==1]
df3$CMI[df3$nyu==1] <- df3$CMI_nyu[df3$nyu==1]

# create CMI deciles 
df3$cmiD <- cut(df3$CMI, breaks=c(quantile(df3$CMI, probs = seq(0, 1, by = 0.1), na.rm=TRUE)),
  labels=c("0-10","10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90","90-100"), include.lowest=TRUE)
table(df3$cmiD)
```

Sample restriction: to teaching hospitals, not govt owned, to large hosps
```{r}
desc <- ddply(df3, .(provid), summarise, Mbeds = mean(beds, na.rm=T), teachever = max(teaching, na.rm=T), own_gvever =  max(own_gv, na.rm=T)) # NYU: 330214  515.5652

ggplot(desc, aes(Mbeds)) +
  geom_histogram() +
  labs(x="Mean number of beds per hospital",
       y="Count of hospitals",
       title="Distribution of hospitals")
sum(desc$Mbeds < 325, na.rm=T)
desc[desc$Mbeds < 325,]

df3 <- merge(df3, desc, by="provid")

# drop hospitals with Mbeds < 200 | Mcases < 1000
df3 <- subset(df3, Mbeds >= 325)

# restrict to teaching hospitals
df3 <- subset(df3, teachever==1)

# drop government owned hospitals
df3 <- subset(df3, own_gvever==0)

table(df3$qt)

# create a balanced panel
length(table(df3$qt))
test <- ddply(df3, .(provid), summarise, Nqtr=sum(!is.na(mdcost_cpi)))
table(test$Nqtr) # 61 hospitals stayed throughout out of 64

myvars <- names(df3) %in% c("Nqtr.x", "Nqtr.y", "Nqtr")
df3 <- df3[!myvars]
df3 <- merge(test,df3, by=c("provid"), all.y=T)
df3 <- subset(df3, Nqtr==length(table(df3$qt)))
table(df3$qt)
```



Trajectory of raw, inflation & CMI adjsuted mean direct costs for NYU & other teaching hospitals separately
```{r}
# # suppress raw data for FY 2013 only for NYU hospitals
df3$mdcost_cpi_cmi <- df3$mdcost_cpi / df3$CMI 
df3$mtcost_cpi_cmi <- df3$mtcost_cpi / df3$CMI 
# df3$mdcost_nyu_cpi_cmi <- df3$mdcost_nyu_cpi / df3$CMI 
# df3$mtcost_nyu_cpi_cmi <- df3$mtcost_nyu_cpi / df3$CMI 

# line graph of raw, just inflation adjusted, and CMI & inflation adjusted cost at the quarter level just for NYU 
library(reshape2)
df4 <- subset(df3, nyu==1)
keep <- c("mdcost", "mdcost_cpi", "mdcost_cpi_cmi", "provid", "qtr")
test <- ddply(df4[keep], .(qtr), numcolwise(mean, na.rm=TRUE))
test <- melt(test[c("mdcost", "mdcost_cpi", "mdcost_cpi_cmi", "qtr")], id.var="qtr")
test$variable <- factor(test$variable, labels = c("Raw", "Inflation-adjusted", "Inflation-adjusted / Mean DRG weights"))

# line graph of CMI & inflation adjusted cost at the quarter level just for NYU 
ggplot(test, aes(x=qtr, y=value, colour=variable, group=variable)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Mean direct cost ($)",
       title="Mean direct cost ($) by quarter",
       subtitle="NYU Hospitals") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,20000),
                     breaks=seq(0,20000,2000) ) +
  scale_colour_discrete(guide = guide_legend(ncol=1))
ggsave("~/Dropbox/Research/VBM/results/mdcost_trajectory_all3.pdf", width=20, height=10, units="cm")

# all other teaching hospitals
df4 <- subset(df3, nyu==0)
keep <- c("mdcost", "mdcost_cpi", "mdcost_cpi_cmi", "provid", "qtr", "dcost_cases")
test <- ddply(df4[keep], .(qtr), numcolwise(mean, na.rm=TRUE))
test <- melt(test[c("mdcost", "mdcost_cpi", "mdcost_cpi_cmi", "qtr", "dcost_cases")], id.var="qtr")
test2 <- test

test <- subset(test, variable!="dcost_cases")
test$variable <- factor(test$variable, labels = c("Raw", "Inflation-adjusted", "Inflation-adjusted / Mean DRG weights"))

# line graph of CMI & inflation adjusted cost at the quarter level just for NYU 
ggplot(test, aes(x=qtr, y=value, colour=variable, group=variable)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Mean direct cost ($)",
       title="Mean direct cost ($) by quarter",
       subtitle="All Other Teaching Hospitals",
       caption="Data source: 103 hospitals that had non-zero mean direct cost throughout the sample period"
       ) +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,20000),
                     breaks=seq(0,20000,2000)) +
  scale_colour_discrete(guide = guide_legend(ncol=1))
ggsave("~/Dropbox/Research/VBM/results/mdcost_trajectory_all3_other.pdf", width=20, height=10, units="cm")

# plot # cases
test <- subset(test2, variable=="dcost_cases")

ggplot(test, aes(x=qtr, y=value, group=1)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Number of cases",
       title="Number of cost cases by quarter",
       subtitle="All Other Teaching Hospitals",
       caption="Data source: 103 hospitals that had non-zero mean direct cost throughout the sample period") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,10000),
                     breaks=seq(0,10000,2000)) 
ggsave("~/Dropbox/Research/VBM/results/cost_cases_trajectory_other.pdf", width=20, height=10, units="cm")


# put NYU & other teaching hospitals in the same graph 
keep <- c("mdcost_cpi_cmi","mtcost_cpi_cmi","CMI", "provid", "qtr", "nyu")
test <- ddply(df3[keep], .(nyu,qtr), numcolwise(mean, na.rm=TRUE))
test <- melt(test[c("mdcost_cpi_cmi", "mtcost_cpi_cmi", "qtr", "nyu", "CMI")], id.var=c("nyu","qtr"))
test$variable <- factor(test$variable, labels = c("Inflation-adjusted mean direct costs / Mean DRG weights", "Inflation-adjusted mean total costs / Mean DRG weights", "Mean DRG weights"))
test$nyu <- factor(test$nyu, levels=c(1,0), labels = c("NYU","Other teaching hospitals"))

test2 <- subset(test, variable=="Mean DRG weights")
ggplot(test2, aes(x=qtr, y=value, colour = factor(nyu), group=nyu)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="DRG weights",
       title="Mean DRG weights by quarter",
       subtitle="NYU vs All Other Teaching Hospitals") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,2),
                     breaks=seq(0,2,0.5)) 
ggsave("~/Dropbox/Research/VBM/results/DRGweight_NYU_vs_other.pdf", width=20, height=10, units="cm")

test2 <- subset(test, variable=="Inflation-adjusted mean direct costs / Mean DRG weights")
ggplot(test2, aes(x=qtr, y=value, colour = factor(nyu), group=nyu)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Mean direct cost ($)",
       title="Inflation-adjusted mean direct costs / Mean DRG weights by quarter",
       subtitle="NYU vs All Other Teaching Hospitals") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,20000),
                     breaks=seq(0,20000,2000)) 
ggsave("~/Dropbox/Research/VBM/results/mdcost_NYU_vs_other.pdf", width=20, height=10, units="cm")

test2 <- subset(test, variable=="Inflation-adjusted mean total costs / Mean DRG weights")
ggplot(test2, aes(x=qtr, y=value, colour = factor(nyu), group=nyu)) +
  geom_line() +
  geom_point() +
  labs(x="Quarter", y="Mean total cost ($)",
       title="Inflation-adjusted mean total costs / Mean DRG weights by quarter",
       subtitle="NYU vs All Other Teaching Hospitals") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom", legend.title=element_blank()) +
  geom_vline(aes(xintercept=4), colour="red") +
  scale_y_continuous(limits = c(0,20000),
                     breaks=seq(0,20000,2000)) 
ggsave("~/Dropbox/Research/VBM/results/mtcost_NYU_vs_other.pdf", width=20, height=10, units="cm")

```

Diff-in-diff analysis
```{r}
#re-create size category
df3$size3 <- factor(ifelse(df3$beds < 450, 1, ifelse(df3$beds < 700, 2, 3)), labels=c("large", "very large", "huge"))
table(df3$size3)

df4 <- df3

did <- lm(lnmdcost_cpi ~ int + int*nyu + cmiD + factor(provid),
          data=df4)
summary(did)

pct_chng <- function(x) 100*(exp(x)-1)

comb <- c("int*nyu + int + factor(provid)",
          "int*nyu + int + factor(provid) + cmiD",
          "int*nyu + int + factor(provid) + cmiD + factor(size3)",
          "int*nyu + int + factor(provid) + cmiD + factor(size3) + dissh + urban")

costs <- c("mdcost_cpi", "mtcost_cpi")
costs_lab <- c("Inflation-adjusted mean direct costs", "Inflation-adjusted mean total costs")
init <- 0
n <- length(comb)

for (col in costs) {
  init <- init + 1
  depv <- c(costs_lab[init])
  
  models <- lapply(paste0( rep(col,n),"~", comb), formula)

  result <- lapply(models, FUN = function(x) {
    # lmcoef <- coef(lm(formula = x, data=patUse3, na.action="na.omit"))
    # k <- length(lmcoef)

    glm(formula = x,
              data = df4,
              family=Gamma(link="log"),
              na.action="na.omit")
              # start=c(20, rep(0,k-1)))
  })

  for (i in 1:n) {
    print(summary(result[[i]]))
  }
  stargazer(result, type="html", report = "vcp*",
            t.auto=F, p.auto=F,
            # keep.stat=c("n","adj.rsq","ll","aic"),
            apply.coef = pct_chng,
            dep.var.labels = depv,
            # covariate.labels = c("Time","Intervention","Time After Intervention"),
            out=paste("~/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/Users/kunheekim/Dropbox/Research/VBM/results/", col, ".htm", sep=""))
            # out=paste("/home/hcmg/kunhee/VBM/", col, ".htm", sep=""))
}
```

get quarterly coefficients for the inflation-adjusted mean direct costs
```{r}
# regression on NYU indicator
did.glm <- glm(formula = mdcost_cpi ~ qtr*nyu + factor(provid) + cmiD + dissh + urban + factor(size3),
              data = df4,
              family=Gamma(link="log"),
              na.action="na.omit")
summary(did.glm)

CI <- confint(did.glm)
CI2 <- 100*(exp(CI)-1)
coeff_pc <- 100*(exp(did.glm$coefficients)-1)

didcoef <- cbind(coeff_pc, CI2)
didcoef <- didcoef[87:104,]
didcoef <- as.data.frame(didcoef)

# create qtr var
didcoef$qtr_n <- c(rep(1:4,4),1:2)
didcoef$qtr_yr <- c(rep(2012,3), 2013, rep(2014,4), rep(2015,4), rep(2016,4), rep(2017,2))

# create empty rows for quarters for FY 2013 
qtr_n <- c(4,1:3)
qtr_yr <- c(2012, rep(2013,3))
coeff_pc <- NA
`2.5 %` <- NA
`97.5 %`<- NA
x <- cbind(coeff_pc, `2.5 %`, `97.5 %`, qtr_n, qtr_yr)
didcoef <- rbind(didcoef,x)

library(zoo)
didcoef$qtr <- as.yearqtr(paste(didcoef$qtr_yr, didcoef$qtr_n, sep=" Q"), format = "%Y Q%q")

# order data by quarter
didcoef <- didcoef[order(didcoef$qtr_yr, didcoef$qtr_n),]

# didcoef$id <- 1:NROW(didcoef)
# ggplot(didcoef) +
#   geom_point(aes(x=id, y=coeff_pc)) +
#   geom_errorbar(aes(x=id, ymin=`2.5 %`, ymax=`97.5 %`)) +
#   scale_y_continuous(limits = c(-50,100))

ggplot(didcoef) +
  geom_point(aes(x=qtr, y=coeff_pc)) +
  geom_errorbar(aes(x=qtr, ymin=`2.5 %`, ymax=`97.5 %`)) +
  geom_vline(aes(xintercept=2014.25), colour="red") +
  scale_y_continuous(limits = c(-50,100),
                     minor_breaks=seq(-50,100,10)) +
  scale_x_continuous(breaks=seq(2012, 2017.25, 0.25),
                     labels=paste(didcoef$qtr_yr, didcoef$qtr_n, sep=" Q")) +
  labs(x="Year-Quarter",
       y="Percentage change (%)",
       title="DRG weight adjusted mean direct costs in NYU vs other teaching hospitals by quarter") +
  theme_grey(base_size=11) +
  theme(axis.text.x = element_text(angle=45, vjust = 0.5))
ggsave("~/Dropbox/Research/VBM/results/did_mdcost_qtr.pdf", width=20, height=10, units="cm")
```

get quarterly coefficients for the inflation-adjusted mean total costs
```{r}
# regression on NYU indicator
did.glm2 <- glm(formula = mdcost_cpi ~ qtr*nyu + factor(provid) + cmiD + dissh + urban + factor(size3),
              data = df4,
              family=Gamma(link="log"),
              na.action="na.omit")
summary(did.glm2)

CI2 <- confint(did.glm2)
CI2 <- 100*(exp(CI2)-1)
coeff_pc <- 100*(exp(did.glm2$coefficients)-1)

didcoef2 <- cbind(coeff_pc, CI2)
didcoef2 <- didcoef2[87:NROW(coeff_pc),]
didcoef2 <- as.data.frame(didcoef2)

# create qtr var
didcoef2$qtr_n <- c(rep(1:4,4),1:2)
didcoef2$qtr_yr <- c(rep(2012,3), 2013, rep(2014,4), rep(2015,4), rep(2016,4), rep(2017,2))

# create empty rows for quarters for FY 2013 
qtr_n <- c(4,1:3)
qtr_yr <- c(2012, rep(2013,3))
coeff_pc <- NA
`2.5 %` <- NA
`97.5 %`<- NA
x <- cbind(coeff_pc, `2.5 %`, `97.5 %`, qtr_n, qtr_yr)
didcoef2 <- rbind(didcoef2,x)

library(zoo)
didcoef2$qtr <- as.yearqtr(paste(didcoef2$qtr_yr, didcoef2$qtr_n, sep=" Q"), format = "%Y Q%q")

# order data by quarter
didcoef2 <- didcoef2[order(didcoef2$qtr_yr, didcoef2$qtr_n),]

ggplot(didcoef2) +
  geom_point(aes(x=qtr, y=coeff_pc)) +
  geom_errorbar(aes(x=qtr, ymin=`2.5 %`, ymax=`97.5 %`)) +
  geom_vline(aes(xintercept=2014.25), colour="red") +
  scale_y_continuous(limits = c(-50,100),
                     minor_breaks=seq(-50,100,10)) +
  scale_x_continuous(breaks=seq(2012, 2017.25, 0.25),
                     labels=paste(didcoef2$qtr_yr, didcoef2$qtr_n, sep=" Q")) +
  labs(x="Year-Quarter",
       y="Percentage change (%)",
       title="DRG weight adjusted mean total costs in NYU vs other teaching hospitals by quarter") +
  theme_grey(base_size=11) +
  theme(axis.text.x = element_text(angle=45, vjust = 0.5))
ggsave("~/Dropbox/Research/VBM/results/did_mtcost_qtr.pdf", width=20, height=10, units="cm")
```

